---
output: html_document
editor_options: 
  chunk_output_type: console
---


<pre>

## Callin Switzer
## 17 Nov 2017
## Multilevel model to visualize bees'
## behavior on the artificial pollen system

## Update 30 July 2018 -- investigate proportion of time rewarded

</pre>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rd = "D:/Dropbox/SonicationBehavior"
# "/Users/cswitzer/Dropbox/SonicationBehavior"
knitr::opts_knit$set(root.dir =rd )
```


```{r}
#install packages
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("tidyverse", "reshape2", 'lme4', 'sjPlot', 
              "multcomp", "plyr", "effects")
ipak(packages)

# set ggplot theme
theme_set(theme_classic() + theme(axis.text=element_text(colour="black")))

# define data and figure directories
# mac
#dataDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehData"
#figDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehFigs"

# windows
dataDir <- "D:/Dropbox/SonicationBehavior/SonBehData"
figDir <- "D:/Dropbox/SonicationBehavior/SonBehFigs"

# check if the directory exists
file.info(dataDir)$isdir

print(paste("last run ", Sys.time()))
print(R.version)
```


### Read in data and double check it

```{r}
sl <- read.csv(file.path(dataDir, '01_CombinedTrials_cleaned.csv'))
colnames(sl)


# calculate proportion of rewards by trt
prop.table(xtabs(~ sl$rewTF + sl$trt), margin = 2)
xtabs(~ sl$rewTF + sl$trt)


# make sure hive is a factor
sl$hive = as.factor(sl$hive)
unique(sl$hive)

# number of trials per bee
xtabs(~sl$beeCol)
hist(xtabs(~sl$beeCol))
max(xtabs(~sl$beeCol))
mean(xtabs(~sl$beeCol))
sd(xtabs(~sl$beeCol))


# number of visits for first trial
xtabs(~sl$beeCol[sl$trialNum == 1])
hist(xtabs(~sl$beeCol[sl$trialNum == 1]))
mean(xtabs(~sl$beeCol[sl$trialNum == 1]))
sd(xtabs(~sl$beeCol[sl$trialNum == 1]))



# number of visits for second trial
xtabs(~sl$beeCol[sl$trialNum == 2])


# number of trials per bee and treatment
xtabs(~sl$beeCol +  sl$trt)
colSums(xtabs(~sl$beeCol +  sl$trt))
# total num
sum(colSums(xtabs(~sl$beeCol +  sl$trt)))

xtabs(~sl$beeCol +  sl$IT_imputed)

# hist(unique(sl$IT_imputed[sl$trt2 == "low"]))
# rug(unique(sl$IT_imputed[sl$trt2 == "low"]))
# 
# hist(unique(sl$IT_imputed[sl$trt2 == "high"]))
# rug(unique(sl$IT_imputed[sl$trt2 == "high"]))

# number of trials per bee and treatment
xtabs(~sl$beeCol +  sl$trialNum)

# tabulate num bees in each trt
tdf <- data.frame(xtabs(~sl$beeCol +  sl$trt))
tdf2 <- acast(tdf, sl.beeCol~sl.trt, value.var="Freq")
tdf3 <- as.data.frame(tdf2 > 0)

trtdf<-table(interaction(tdf3$full, tdf3$high, tdf3$low))
names(trtdf) <- c("FULL ONLY", "FULL-HIGH", "FULL-LOW", "NA")
trtdf

# get number of buzzes in each treatment
colSums(tdf2)

# get the number of bees for each treatment (every bee has "full" treatment)
apply(tdf2, MARGIN = 2, function(x) length(x[x>0]))


head(sl)
dim(sl) # should be 24303 rows

# hive 5 is most common
table(sl$hive, useNA = 'always')

## make sure all bee colors are lowercase
sl$beeCol <- tolower(sl$beeCol)

# make sure there are values lower than 220 and higher than 450 
# (the cutoff for buzzes used in the experiment)
hist(sl$freq, breaks = seq(215, 450, by = 10))
nrow(sl[sl$freq < 220 | sl$freq > 450,]) # should have 0 rows


# look at treatments
xtabs(~sl$beeCol+ trt, data = sl )

# find percentage reward by treatment
mean(grepl("[tT]", as.character(sl$rewTF))) # overall mean

# percentage that were rewarded by treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, mean)

# total number of trials for each treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, length)

# total number of rewards per treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, FUN = function(x) sum(x))

# total number of trials that were unrewarded per treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, FUN = function(x) sum(!x))


# calculate mean freq per bee per trial
d1 <- as.data.frame((tapply(sl$freq, INDEX = list(sl$beeCol, sl$trt), mean)))
d2 <- as.data.frame((tapply(sl$freq, INDEX = list(sl$beeCol, sl$trt), length)))
d3 <- data.frame(d1, d2)
d3$beeCol = row.names(d3)


d4 <- melt(d3, id.vars=c("beeCol"), measure.vars=c("full", "high", "low" ), value.name = "buzzFreq")
d5 <- melt(d3, id.vars=c("beeCol"), measure.vars=c("full.1",  "high.1", "low.1" ), value.name = "NumBuzzes")


d6 <- data.frame(d4, d5)
d7 <- d6[!is.na(d6$buzzFreq),]
d7

# this plot shows how individual bees changed
ggplot(d7, aes(x = variable, y = buzzFreq)) + 
  geom_boxplot(aes(fill = variable == "full"), alpha = 0.2) + 
  geom_point(aes(color = variable == "full", size = NumBuzzes)) + 
  geom_line(aes(group = beeCol), alpha = 0.2) + 
  scale_fill_viridis_d(name = "hihi", guide = FALSE, option = "A", end = 0.6) + 
  scale_color_viridis_d(name = "hihi", guide = FALSE, option = "A", end = 0.6) + 
  labs(x = "treatment")



```

### Visualizations and modeling

```{r aggregate, cache = TRUE}
# calculate trial averages and plot
sl$colNum = paste(sl$beeCol, sl$hive, sep = "_")
sl$trt <- as.character(sl$trt)
sl$trt[sl$trt == "full" & sl$trialNum >1 ] <- "full_2"


aggdata <- aggregate(sl$freq, by=list(colNum = sl$colNum,trialNum =sl$trialNum, trt = sl$trt), FUN=mean, na.rm=TRUE)
colnames(aggdata)[colnames(aggdata) == "x"] = "freq"

aggdata_sd <- aggregate(sl$freq, by=list(colNum = sl$colNum,trialNum =sl$trialNum, trt = sl$trt), FUN=sd, na.rm=TRUE)
colnames(aggdata_sd)[colnames(aggdata_sd) == "x"] = "freq_sd"

aggdata = merge(aggdata, aggdata_sd)
#aggdata$trt = sapply(1:nrow(aggdata), FUN = function(x){sl[sl$colNum == aggdata$colNum[x] & sl$trialNum == aggdata$trialNum[x], "trt"][1]})
aggdata = aggdata[order(aggdata$colNum, aggdata$trialNum, decreasing = FALSE), ]
agg_sm = aggdata[aggdata$trialNum <= 2, ]
rownames(agg_sm) = 1:nrow(agg_sm)
agg_sm



agg_sm[duplicated(data.frame(agg_sm$colNum, agg_sm$trialNum)), ]






diffdf <- sapply(unique(agg_sm$colNum), FUN = function(x){
  tmp = agg_sm[agg_sm$colNum == x, ]
  if(nrow(tmp) <= 1)
    diff = NA
  else
    diff = tmp$freq[tmp$trialNum == 2] - tmp$freq[tmp$trialNum == 1]
  return(diff)
})

trtDF = sapply(unique(agg_sm$colNum), FUN = function(x){
  tmp = agg_sm[agg_sm$colNum == x, ]
  ttrs = paste(tmp$trt[tmp$trialNum == 1], tmp$trt[tmp$trialNum == 2], sep = "_")
  return(ttrs)
})

buzzdiffs = data.frame(trtDF, diffdf)

ggplot(buzzdiffs, aes(x = trtDF, y= diffdf)) + 
  geom_boxplot() + 
  geom_point()

         
agg2 = aggregate(sl$freq, by=list(colNum = sl$colNum, fullTrt = sl$trt == "full", trt = sl$trt), FUN=mean, na.rm=TRUE)
colnames(agg2)[colnames(agg2) == "x"] = "freq"
agg2$trt = as.character(agg2$trt)

diffdf <- t(as.data.frame(t(sapply(unique(agg2$colNum), FUN = function(x){
  tmp = agg2[agg2$colNum == x, ]
  if(nrow(tmp) <= 1)
    return(NA)
  if (length(unique(tmp$trt)) > 2){
    tmp = tmp[tmp$trt != "full_2", ]
  }
  diff =  tmp$freq[!tmp$fullTrt] - tmp$freq[tmp$fullTrt]
  return(diff)

}))))

length(diffdf)

trtDF = sapply(unique(agg2$colNum), FUN = function(x){
  tmp = agg2[agg2$colNum == x, ]
  if (length(unique(tmp$trt)) > 2){
    tmp = tmp[tmp$trt != "full_2", ]
  }
  ttrs = paste(tmp$trt[tmp$fullTrt], tmp$trt[!tmp$fullTrt], sep = "_")
  return(ttrs)
})

length(trtDF)

buzzdiffs = data.frame(trtDF, diffdf)
tapply(buzzdiffs$diffdf, INDEX = buzzdiffs$trtDF, mean)

ggplot(droplevels(buzzdiffs[buzzdiffs$trtDF != "full_", ]), aes(x = trtDF, y= as.numeric(diffdf))) + 
  geom_boxplot() + 
  geom_point() + 
  labs(y = "diff in frequency -- averages")

```



### MODELING


```{r}

sl$trt = relevel(factor(sl$trt), ref = "full")


sl$trt2 = mapvalues(sl$trt, from = c("full", "full_2", "high", "low"), 
                    to = c("full", "full", "high", "low"))
# summary for paper
# fit a varying slope and intercept for colNum (bee ID), and allow the slope of the trialNum
# variable to vary by colNum (beeID)
m2 = lmer(freq ~ trt2 + hive + trialNum + IT_imputed + (1+trialNum|colNum), data = sl, REML = FALSE)

# this model estimates a global intercept
# random effect intercept for colNum (beeID)
# a single global estimate for trialNum
# the effect of trialNum within each level of colNum
# the correlation between intercept of trialNum across levels of colNum


vif.mer <- function (fit) {
    ## adapted from rms::vif

    v <- vcov(fit)
    nam <- names(fixef(fit))

    ## exclude intercepts
    ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
    if (ns > 0) {
        v <- v[-(1:ns), -(1:ns), drop = FALSE]
        nam <- nam[-(1:ns)]
    }

    d <- diag(v)^0.5
    v <- diag(solve(v/(d %o% d)))
    names(v) <- nam
    v
}

vif.mer(m2) # fine

m2_1 = lmer(freq ~ trt2* IT_imputed + hive + trialNum  + (1+trialNum|colNum), data = sl, REML = FALSE)

# compare models with BIC
BIC(m2, m2_1) # m2 (no interaction) is better
anova(m2, m2_1) # Note: this disagrees with BIC



m3 = update(m2, .~. - hive)
BIC(m2, m3) # m3 is better (without hive)
anova(m2, m3) # Note: disagrees with BIC

# refit best model with REML = TRUE
m3 <- update(m3, .~., REML = TRUE)
summary(m3) # summary for paper -- leave trial number in, to account for random effect
plot(m3)
qqnorm(ranef(m3)$colNum[[1]])
qqline(ranef(m3)$colNum[[1]])

sjp.lmer(m3) # plot random effects to find any outliers
sjp.lmer(m3, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects


# post-hoc tests -- pvals for paper
summary(glht(m3, linfct = mcp(trt2 = "Tukey")), test = adjusted("none"))

## bonf adjusted pvals
summary(glht(m3, linfct = mcp(trt2 = "Tukey")), test = adjusted("bonf"))
```


### Generate CI's
```{r bootstrap, cache = TRUE}

# set number of bOotstrap samples
nbootSims = 10

ITmean = mean(tapply(sl$IT_imputed, INDEX = sl$beeCol, FUN = function(x) x[1] ))
print(ITmean)

# don't need hive, because that's not in the model we chose (above)
pframe <- data.frame(trt2 = levels(droplevels(sl$trt2)), IT_imputed = ITmean,  colNum = 99999, trialNum = 2)
pframe$freq <- 0
pp <- predict(m3, newdata = pframe, re.form=NA, type = 'response') # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(m3, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<-bb2_se[1,]
pframe$bhi<-bb2_se[2,]
pframe$predMean <- pp
pframe <- pframe[, c('trt2', "blo", "bhi", "predMean")]
pframe
```

### Make plots for paper

```{r}
#plot 95% confidence intervals
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
# rename labels for plot
pframe$trt3 = mapvalues(pframe$trt2, from = c("full", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", "High range\n(340 - 390 Hz)", "Low range\n(220 - 330 Hz)"))

pframe
g0 <- ggplot(pframe, aes(x=trt3, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication frequency (Hz)", x = "Frequency range for reward") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)+
  theme(plot.background = element_rect(fill = "transparent",colour = NA), 
        panel.background = element_rect(fill = "transparent",colour = NA)) + 
  annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 360, label=c("a", "b", "c"),
                color="black") 
g0

ggsave(plot = g0, filename = file.path(figDir, "SonicationFreqPredsAndCI_unadjusted.png"), width = 5, height = 4)

# save with transparent background
g0 <- ggplot(pframe, aes(x=trt3, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication frequency (Hz)", x = "Frequency range for reward") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)+
  theme(plot.background = element_rect(fill = "transparent",colour = NA), 
        panel.background = element_rect(fill = "transparent",colour = NA)) + 
  annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 360, label=c("a", "b", "b"),
                color="black") 
g0

ggsave(plot = g0, filename = file.path(figDir, "SonicationFreqPredsAndCI_adjustedPvals.pdf"), width = 5, height = 4)




# use the "effects" library to get CI's
# this basically gives the same result as bootstrapping
# it's a good double check
ee <- as.data.frame(Effect(c("trt2"), 
                           fixed.predictors =list(given.values = c(trialNum = 2, 
                                                                   IT_imputed = ITmean)),  
                           m3) )


# compare two methods -- very similar
ee
pframe

g1 <- ggplot(ee, aes(x=trt2, y=fit))+
     geom_point()+ 
     labs(y = "Sonication Frequency (Hz)", x = "Treatment") + 
     geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1)+
     theme_classic() + 
    annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 360, label=c("a", "b", "c"),
                color="black") 
g1

```

### ____ END OF ANALYSIS FOR PAPER ____ ### 



### Analysis for amplitude

```{r ampAnalysis}

# refref: here try log amp_acc and REML = FALSE with BIC

# interaction model
# note log-transformation to make model fit assumptions better
maa0 = lmer(log(amp_acc) ~ trt2* IT_imputed + hive + trialNum  + (1+trialNum|colNum), data = sl, REML = FALSE)

# main effect model
maa1 = lmer(log(amp_acc) ~ trt2 + hive + trialNum + IT_imputed + (1+trialNum|colNum), data = sl, REML = FALSE)
BIC(maa0, maa1) # use no interaction (keep maa1)
anova(maa0, maa1) # note that this agrees with likelihood ratio test


maa2 = update(maa1, .~. - trt2)
BIC(maa1, maa2) # keep treatment (maa1)
anova(maa1, maa2) # agrees with LRT; p-val for paper, if needed

maa3 <- update(maa1, .~. - hive)
BIC(maa1, maa3) # remove hive (maa3)
anova(maa1, maa3) # agrees with LRT


# refit with REML = TRUE for paper
maa3 <- update(maa3, .~., REML = TRUE)
summary(maa3)


# diagnostics
plot(maa3)
qqnorm(ranef(maa1)$colNum[[1]])
qqline(ranef(maa1)$colNum[[1]])


sjp.lmer(maa3) # plot random effects to find any outliers
sjp.lmer(maa3, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects


# post-hoc tests -- pvals for paper
summary(glht(maa3, linfct = mcp(trt2 = "Tukey")), test = adjusted("none"))


# post-hoc tests with bonf adjustment
summary(glht(maa3, linfct = mcp(trt2 = "Tukey")), test = adjusted("bonf"))

```


### Generate CI's for amplitude

```{r bootstrapAmp, cache = TRUE}
# set number of bootstrap samples
nbootSims2 = 10

# don't need to include hive, but will set trialNum to 2
pframe <- data.frame(trt2 = levels(droplevels(sl$trt2)), 
                     IT_imputed = ITmean,  colNum = 99999, trialNum = 2)
pframe$amp_acc <- 0
# exponentiate to get back to original scale
pp <- exp(predict(maa3, newdata = pframe, re.form=NA, type = 'response')) # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(maa3, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims2)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))

#exponentiate to get back to orignal scale
pframe$blo<-exp(bb2_se[1,])
pframe$bhi<-exp(bb2_se[2,])
pframe$predMean <- pp
pframe <- pframe[, c('trt2', "blo", "bhi", "predMean")]
pframe
```


### Make plots for amplitude for paper

```{r amplitudePlots}
#plot 95% confidence intervals
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
pframe$trt3 = mapvalues(pframe$trt2, from = c("full", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", "High range\n(340 - 390 Hz)", "Low range\n(220 - 330 Hz)"))

pframe
ga0 <- ggplot(pframe, aes(x=trt3, y=predMean))+
     geom_point()+ 
     labs( y = expression ("Sonication amplitude "(m~s^{-2})), x = "Frequency range for reward") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)+
     theme(axis.text.x = element_text(angle = 45, hjust = 1), 
           legend.position = 'none') +
     theme_classic() + 
  annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 58, label=c("a", "b", "c"),
                color="black") 
ga0
ggsave(plot = ga0, filename = file.path(figDir, "SonicationAmpPredsAndCI_unadjusted.pdf"), width = 5, height = 4)


# make plot using bonferroni adjustments

ga0 <- ggplot(pframe, aes(x=trt3, y=predMean))+
     geom_point()+ 
     labs( y = expression ("Sonication amplitude "(m~s^{-2})), x = "Frequency range for reward") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)+
     theme(axis.text.x = element_text(angle = 45, hjust = 1), 
           legend.position = 'none') +
     theme_classic() + 
  annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 58, label=c("a", "b", "a"),
                color="black") 
ga0
ggsave(plot = ga0, filename = file.path(figDir, "SonicationAmpPredsAndCI_BonfAdjusted.pdf"), width = 5, height = 4)

```


# Visualize aggregated data for amplitude

```{r}

aggdata <- aggregate(sl$amp_acc, by=list(colNum = sl$colNum,trialNum =sl$trialNum, trt = sl$trt), FUN=mean, na.rm=TRUE)
colnames(aggdata)[colnames(aggdata) == "x"] = "amp"

aggdata_sd <- aggregate(sl$amp_acc, by=list(colNum = sl$colNum,trialNum =sl$trialNum, trt = sl$trt), FUN=sd, na.rm=TRUE)
colnames(aggdata_sd)[colnames(aggdata_sd) == "x"] = "amp_sd"

aggdata = merge(aggdata, aggdata_sd)
aggdata = aggdata[order(aggdata$colNum, aggdata$trialNum, decreasing = FALSE), ]
agg_sm = aggdata[aggdata$trialNum <= 2, ]
rownames(agg_sm) = 1:nrow(agg_sm)
agg_sm



agg_sm[duplicated(data.frame(agg_sm$colNum, agg_sm$trialNum)), ]

ggplot(agg_sm, aes(x = trt, y = amp, fill = trialNum > 1)) +
  geom_boxplot(alpha = 0.2) + 
  geom_point(aes(color = trialNum>1)) + 
  geom_line(aes(group = colNum))


diffdf <- sapply(unique(agg_sm$colNum), FUN = function(x){
  tmp = agg_sm[agg_sm$colNum == x, ]
  if(nrow(tmp) <= 1)
    diff = NA
  else
    diff = tmp$amp[tmp$trialNum == 2] - tmp$amp[tmp$trialNum == 1]
  return(diff)
})

trtDF = sapply(unique(agg_sm$colNum), FUN = function(x){
  tmp = agg_sm[agg_sm$colNum == x, ]
  ttrs = paste(tmp$trt[tmp$trialNum == 1], tmp$trt[tmp$trialNum == 2], sep = "_")
  return(ttrs)
})

buzzdiffs = data.frame(trtDF, diffdf)

ggplot(buzzdiffs, aes(x = trtDF, y= diffdf)) + 
  geom_boxplot() + 
  geom_point()+ 
  labs(y = "amplitude difference m/s/s")

         
agg2 = aggregate(sl$amp_acc, by=list(colNum = sl$colNum, fullTrt = sl$trt == "full", trt = sl$trt), FUN=mean, na.rm=TRUE)
colnames(agg2)[colnames(agg2) == "x"] = "amp"
agg2$trt = as.character(agg2$trt)

diffdf <- t(as.data.frame(t(sapply(unique(agg2$colNum), FUN = function(x){
  tmp = agg2[agg2$colNum == x, ]
  if(nrow(tmp) <= 1)
    return(NA)
  if (length(unique(tmp$trt)) > 2){
    tmp = tmp[tmp$trt != "full_2", ]
  }
  diff =  tmp$amp[!tmp$fullTrt] - tmp$amp[tmp$fullTrt]
  return(diff)

}))))

length(diffdf)

trtDF = sapply(unique(agg2$colNum), FUN = function(x){
  tmp = agg2[agg2$colNum == x, ]
  if (length(unique(tmp$trt)) > 2){
    tmp = tmp[tmp$trt != "full_2", ]
  }
  ttrs = paste(tmp$trt[tmp$fullTrt], tmp$trt[!tmp$fullTrt], sep = "_")
  return(ttrs)
})

length(trtDF)

buzzdiffs = data.frame(trtDF, diffdf)
tapply(buzzdiffs$diffdf, INDEX = buzzdiffs$trtDF, mean)

ggplot(droplevels(buzzdiffs[buzzdiffs$trtDF != "full_", ]), aes(x = trtDF, y= as.numeric(diffdf))) + 
  geom_boxplot() + 
  geom_point()+ 
  labs(y = "amplitude difference m/s/s")
```


## Plot amplitude vs. frequency for initial trial, accounting for size

```{r}
ggplot(sl[sl$trialNum == 1, ], aes(x = freq, y = amp_acc)) + 
  geom_point(position = position_jitter(width = 3), alpha = 0.2) + 
  theme_classic() + 
  geom_smooth(aes(group = trt2)) + 
  labs( y = expression ("Sonication amplitude "(m~s^{-2})), x = "Frequency (Hz)")

centerFreq = scale(sl[sl$trialNum == 1, "freq"])

m1 <- lmer(log(amp_acc)~centerFreq + I(centerFreq^2) + I(centerFreq^3) + IT_imputed + (1|beeCol), data = sl[sl$trialNum == 1, ])
summary(m1)

plot(m1)

# predict -- using mean IT span
ppdf <- data.frame(centerFreq = sort(unique(centerFreq)), IT_imputed = ITmean, beeCol = 99999, acc_amp = 0)
# exponentiate to get back to original scale
ppdf$predAmp = exp(predict(m1, newdata = ppdf, type = "response", re.form = NA))



cf_Unscaled = ppdf$centerFreq * attr(centerFreq, 'scaled:scale') + attr(centerFreq, 'scaled:center')
ggplot(ppdf, aes(x = cf_Unscaled, y = predAmp))+ 
  geom_line() + 
  labs(x = "Sonication Frequency (Hz)", y = expression ("Sonication amplitude "(m~s^{-2})) ) + 
  geom_point(data = sl[sl$trialNum == 1, ], 
             aes(x = freq, y = amp_acc), 
             alpha = 0.2, position = position_jitter(width =2), pch =16, size = .5) 
ggsave(filename = file.path(figDir, "FreqVsAmp_1stTrial_rawData.pdf"), width = 4, height = 3)

g22 <- ggplot(ppdf, aes(x = cf_Unscaled, y = predAmp))+ 
  geom_line() + 
  labs(x = "Sonication Frequency (Hz)", y = expression ("Predicted sonication amplitude "(m~s^{-2})) ) 
g22
ggsave(filename = file.path(figDir, "FreqVsAmp_1stTrial.pdf"), width = 4, height = 3)
  
```


## analyze time between buzzes
```{r}
# convert time to datetime format
options(digits.secs=6)
sl$dateTime_fmt= strptime(x = sl$datetime_str, format = "%Y-%m-%d  %H:%M:%OS" )

sl %>% head

# calculate time since beginning of trial
timeSinceStart = list()
timeDiff = list()
timeSinceReward = list()
IDS = character()
prevRewardTime = NA
buzzesSinceReward = list()

# calculate num of buzzes since last reward


for(ii in 1:nrow(sl)){
  
  tmp = sl[ii, c("BeeNumCol", "dateTime_fmt", "rewTF", "index")]
  
  if(!(tmp$BeeNumCol %in% IDS)){
    timeSinceStart[ii] = difftime(tmp$dateTime_fmt, tmp$dateTime_fmt, units = "min")
    timeDiff[ii] = difftime(tmp$dateTime_fmt, tmp$dateTime_fmt, units = "min")
    timeSinceReward[ii] = difftime(tmp$dateTime_fmt, tmp$dateTime_fmt, units = "min")
    buzzesSinceReward[ii] = NA
    if(as.character(tmp$rewTF) == " T"){
      prevRewardTime = tmp$dateTime_fmt
      prevRewardBuzzNum = tmp$index
    }
    else{
       prevRewardTime = NA
       prevRewardBuzzNum = NA
    }
    startTime = tmp$dateTime_fmt
    IDS = append(IDS, as.character(tmp$BeeNumCol))
  }
  else{
    timeSinceStart[ii] = difftime(tmp$dateTime_fmt, startTime, units = "min")
    timeDiff[ii] = difftime(tmp$dateTime_fmt, prevTime, units = "min")
    
    timeSinceReward[ii] = difftime(tmp$dateTime_fmt, prevRewardTime, units = "min")
    buzzesSinceReward[ii] = tmp$index - prevRewardBuzzNum
    if(tmp$rewTF == " T"){
      prevRewardTime = tmp$dateTime_fmt
      prevRewardBuzzNum = tmp$index
    }
  }
  prevTime = tmp$dateTime_fmt

}


sl$timeSinceStart = unlist(timeSinceStart)
sl$timeSinceLastBuzz = unlist(timeDiff)
sl$timeSinceReward = unlist(timeSinceReward)
sl$buzzesSinceReward = unlist(buzzesSinceReward)

#plot(sl$index)
sl %>% head

plot(sl$timeSinceStart[1:500])
plot(sl$timeSinceLastBuzz[1:1000])

plot(sl$buzzesSinceReward[1:5000])

ggplot(data = sl[1:10000, ], mapping = aes(x = index, y = timeSinceReward, color= trt)) + 
# geom_boxplot() + 
   geom_point() + 
  facet_wrap(~trt) + 
  scale_y_log10() + 
  theme(legend.position = "none") + 
  geom_smooth(color = 'grey40') + 
  scale_color_viridis_d(option = "magma", end = 0.9)

ggplot(data = sl[1:10000, ], mapping = aes(x = index, y = timeSinceReward, color= trt)) + 
# geom_boxplot() + 
   geom_point() + 
  facet_wrap(~trt) + 
  scale_y_log10() + 
  theme(legend.position = "none") + 
  geom_smooth(color = 'grey40') + 
  scale_color_viridis_d(option = "magma", end = 0.9)



# plot time vs. reward 

ggplot(data = sl[1000:2000, ], mapping = aes(x = timeSinceStart, y = timeSinceReward, color= trt)) + 
# geom_boxplot() + 
   geom_point(alpha = 0.5) + 
  geom_line() + 
  facet_wrap(~BeeNumCol, scales = 'free') + 
  ylim(-0.1, 1.1) + 
  #geom_smooth() + 
  #theme(legend.position = "none") + 
  scale_color_viridis_d(option = "magma", end = 0.9, begin = 0.2,direction = -1)

# plot buzzes b/w rewards
ggplot(data = sl[grepl(pattern = "t", ignore.case = TRUE, x = sl$rewTF) & !(sl$trt %in% c("full", "full_2")), ][1:5000,], mapping = aes(x = index, y = buzzesSinceReward, color= trt)) + 
# geom_boxplot() + 
   geom_point(alpha = 0.5) + 
  geom_line() + 
  facet_wrap(~BeeNumCol, scales = 'free_x') + 
  #ylim(-0.1, 1.1) + 
  #geom_smooth() + 
  #theme(legend.position = "none") + 
  scale_color_viridis_d(option = "magma", end = 0.9, begin = 0.2,direction = -1)




unique(sl[1000:5000, ]$BeeNumCol)

View(sl[sl$BeeNumCol == "Beelimered11_16Dec2016_Hive5_low", ])
View(sl[grepl(pattern = "redpink1", x = as.character(sl$BeeNumCol)), ])
ggplot(data = sl[sl$BeeNumCol == "Beelimered11_16Dec2016_Hive5_low", ], mapping = aes(x = index, y = buzzesSinceReward, color= trt)) + 
# geom_boxplot() + 
   geom_point(alpha = 0.5) + 
  geom_line() + 
  facet_wrap(~BeeNumCol, scales = 'free') + 
  #ylim(-0.1, 1.1) + 
  #geom_smooth() + 
  #theme(legend.position = "none") + 
  scale_color_viridis_d(option = "magma", end = 0.9, begin = 0.2,direction = -1)


ggplot(data = sl[grepl(pattern = "redpink", x = as.character(sl$BeeNumCol)), ], mapping = aes(x = index, y = buzzesSinceReward, color= trt)) + 
# geom_boxplot() + 
   geom_point(alpha = 0.5) + 
  geom_line() + 
  facet_wrap(~BeeNumCol, scales = 'free') + 
  #ylim(-0.1, 1.1) + 
  #geom_smooth() + 
  #theme(legend.position = "none") + 
  scale_color_viridis_d(option = "magma", end = 0.9, begin = 0.2,direction = -1)

ggplot(sl[, ], aes(x = index, y = grepl("t", x = rewTF, ignore.case = TRUE)*1)) +
   geom_point() + 
  geom_smooth() + 
facet_wrap(~trt, scales = 'free')

binomial_smooth <- function(...){
  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
}


ggplot(sl[, ], aes(x = timeSinceStart, y = grepl("t", x = rewTF, ignore.case = TRUE)*1)) +
   geom_point() + 
  binomial_smooth() +  
facet_wrap(~trt, scales = 'free')

ggplot(sl[sl$trt2 %in% c("low", "high"), ], aes(x = index, y = grepl("t", x = rewTF, ignore.case = TRUE)*1)) +
   geom_point() + 
  binomial_smooth(aes(color = as.factor(trialNum)), se = FALSE) +  
facet_wrap(~interaction(beeCol , hive, trt), scales = 'free') + 
  scale_color_viridis_d(end = 0.9, option = "magma", direction = -1)

# refref: visualize using overall buzz number (not reset for each trial), but still
# color by trial

IDS = character()
sl <- sl[order(sl$dateTime_fmt), ]
ID1 = interaction(sl$beeCol, sl$hive, sl$trt2)
buzzctr = 0
overallBuzz = numeric(length = nrow(sl))
for(ii in 1:nrow(sl)){
    tmp = ID1[ii]
    if(tmp %in% IDS){
      buzzctr =  max(overallBuzz[ID1 == tmp]) + 1
    }
    else{
      buzzctr = 0
      IDS = append(IDS, as.character(tmp))
    }
    overallBuzz[ii] = buzzctr
    print(ii)
}
sl$overallBuzz = overallBuzz
    
ggplot(sl[sl$trt2 != "full", ], aes(x = overallBuzz, 
                   y = grepl("t", x = rewTF, ignore.case = TRUE)*1, 
                   color = interaction(beeCol, hive, trt2, trialNum))) +
   geom_point() + 
  geom_smooth(se = FALSE) +  
  theme(legend.position = "none") + 
facet_wrap(~interaction(beeCol, hive, trt2), scales = 'free')




ggplot(sl[sl$trt2 != "full", ], aes(x = overallBuzz, 
                   y = freq, 
                   color = grepl("t", x = rewTF, ignore.case = TRUE))) +
   geom_point() + 
  stat_smooth(method = "loess", se = FALSE, color = "grey40", span = 0.9, lwd= 5) +  
  #theme(legend.position = "none") + 
facet_wrap(~interaction(beeCol, hive, trt2), scales = 'free') + 
  scale_color_viridis_d(name = "Reward")


ggplot(sl[sl$trt2 != "full", ], aes(x = overallBuzz, 
                   y = freq)) +
  geom_ribbon(aes(ymin=lowFrq,ymax=highFrq), fill="grey10", alpha= 0.9, color = NA) +
   geom_hex(binwidth = c(150, 10)) + 
  
  #theme(legend.position = "none") + 
facet_wrap(interaction(beeCol, hive, trt2)~., ncol = 6) + 
  scale_fill_viridis_c(name = "Num buzzes") + 
  geom_hline(aes(yintercept = lowFrq), color = 'grey40', size = 1, lty = 2) + 
  geom_hline(aes(yintercept = highFrq), color = 'grey40', size = 1, lty = 2) 
  


# refref calculate time differences b/w rewards
ggplot(sl[grepl(pattern = "t", x = as.character(sl$rewTF), ignore.case = TRUE) & sl$trt != "full", ], aes(x = as.factor(trialNum), y = timeSinceReward)) + 
         geom_point(aes(color= as.factor(trialNum))) + 
  facet_wrap(interaction(beeCol, hive, trt2)~., ncol = 6, scales = 'free_x') +
  ylim(c(0,1))


# num buzzes b/w rewards
ggplot(sl[grepl(pattern = "t", x = as.character(sl$rewTF), ignore.case = TRUE) & sl$trt2 != "full", ], aes(x = overallBuzz, y = buzzesSinceReward)) + 
         geom_point() + 
  facet_wrap(interaction(beeCol, hive, trt2)~., ncol = 6)


# num buzzes b/w rewards
ggplot(sl[grepl(pattern = "t", x = as.character(sl$rewTF), ignore.case = TRUE) & sl$trt2 != "full", ], aes(x = interaction(index < 50, as.factor(trialNum)), y = buzzesSinceReward, fill = index < 50)) + 
         geom_boxplot() + 
  
  theme(axis.text.x = element_text(angle = 90)) + 
  facet_wrap(interaction(beeCol, hive, trt2)~., ncol = 6, scales = 'free_x') + 
  scale_color_viridis_d()


#refref: calculate continuous time for each bee
sl$continuousTime = 0
sl$beeID = interaction(sl$beeCol, sl$hive)
jj = "orange.3"
kk = 1
for(jj in unique(sl$beeID)){
  tmp = sl[sl$beeID == jj, ]
  tmp %>% head()
  # sort
  tmp <- tmp[order(tmp$dateTime_fmt), ]
  
  times = numeric()
  trials = numeric()
  kk = 1
  for(kk in 1:nrow(tmp)){
    if(!(tmp$trialNum[kk] %in% trials)){
      trials <<- append(trials, tmp$trialNum[kk])
        if(kk == 1){
        times[kk] = 0
        }
      else{
        times[kk-1] = max(times)
      }
      
    }
    if(kk == 1){
        times[kk] = 0
        }
    else{
      times[kk] =  tmp$timeSinceLastBuzz[kk] + times[kk -1]
    }
  }
  tmp$continuousTime = times
  #plot(times)
  sl <- merge(sl, tmp, all = TRUE)
  
}

sl %>% head()



# proportion of rewarded buzzes over time
ggplot(sl, aes(x = timeSinceStart, y = grepl("t", x = rewTF, ignore.case = TRUE)* 1)) + 
  geom_point() + 
  theme(legend.position = "none") + 
  stat_smooth(method = "loess", aes(color = interaction(beeCol, hive, trt2, trialNum)), se = FALSE) + 
  facet_wrap(interaction(beeCol, hive, trt2)~., ncol = 6, scales = 'free_x') + 
  scale_color_viridis_d() + ylim(c(0, 1))

```


### analyze the change in num of buzzes b/w rewards to see if they decrease over time

```{r}




```

```{r}
sessionInfo()
```

