---
output: html_document
editor_options: 
  chunk_output_type: console
---


<pre>

## Callin Switzer
## 17 Nov 2017
## Multilevel model to visualize bees'
## behavior on the artificial pollen system

</pre>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/cswitzer/Dropbox/SonicationBehavior")
```


```{r}
#install packages
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("ggplot2", "reshape2", 'lme4', 'sjPlot', "multcomp", "plyr", "effects")
ipak(packages)

# set ggplot theme
theme_set(theme_classic() + theme(axis.text=element_text(colour="black")))

# define data and figure directories
dataDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehData"
figDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehFigs"

print(paste("last run ", Sys.time()))
print(R.version)
```


### Read in data and double check it

```{r}
sl <- read.csv(file.path(dataDir, '01_CombinedTrials_cleaned.csv'))

# make sure hive is a factor
sl$hive = as.factor(sl$hive)

head(sl)
dim(sl) # should be 24303 rows

# hive 5 is most common
table(sl$hive, useNA = 'always')

## make sure all bee colors are lowercase
sl$beeCol <- tolower(sl$beeCol)

# make sure there are values lower than 220 and higher than 450 
# (the cutoff for buzzes used in the experiment)
hist(sl$freq, breaks = seq(215, 450, by = 10))
nrow(sl[sl$freq < 220 | sl$freq > 450,]) # should have 0 rows


# look at treatments
xtabs(~sl$beeCol+ trt, data = sl )

# find percentage reward by treatment
mean(grepl("[tT]", as.character(sl$rewTF))) # overall mean

# percentage that were rewarded by treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, mean)

# total number of trials for each treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, length)

# total number of rewards per treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, FUN = function(x) sum(x))

# total number of trials that were unrewarded per treatment
tapply((grepl("[tT]", as.character(sl$rewTF))), INDEX= sl$trt, FUN = function(x) sum(!x))
```

### Visualizations and modeling

```{r aggregate, cache = TRUE}
# calculate trial averages and plot
sl$colNum = paste(sl$beeCol, sl$hive, sep = "_")
sl$trt <- as.character(sl$trt)
sl$trt[sl$trt == "full" & sl$trialNum >1 ] <- "full_2"


aggdata <- aggregate(sl$freq, by=list(colNum = sl$colNum,trialNum =sl$trialNum, trt = sl$trt), FUN=mean, na.rm=TRUE)
colnames(aggdata)[colnames(aggdata) == "x"] = "freq"

aggdata_sd <- aggregate(sl$freq, by=list(colNum = sl$colNum,trialNum =sl$trialNum, trt = sl$trt), FUN=sd, na.rm=TRUE)
colnames(aggdata_sd)[colnames(aggdata_sd) == "x"] = "freq_sd"

aggdata = merge(aggdata, aggdata_sd)
#aggdata$trt = sapply(1:nrow(aggdata), FUN = function(x){sl[sl$colNum == aggdata$colNum[x] & sl$trialNum == aggdata$trialNum[x], "trt"][1]})
aggdata = aggdata[order(aggdata$colNum, aggdata$trialNum, decreasing = FALSE), ]
agg_sm = aggdata[aggdata$trialNum <= 2, ]
rownames(agg_sm) = 1:nrow(agg_sm)
agg_sm



agg_sm[duplicated(data.frame(agg_sm$colNum, agg_sm$trialNum)), ]

ggplot(agg_sm, aes(x = trt, y = freq, fill = trialNum > 1)) +
  geom_boxplot(alpha = 0.2) + 
  geom_point(aes(color = trialNum>1)) + 
  geom_line(aes(group = colNum))


diffdf <- sapply(unique(agg_sm$colNum), FUN = function(x){
  tmp = agg_sm[agg_sm$colNum == x, ]
  if(nrow(tmp) <= 1)
    diff = NA
  else
    diff = tmp$freq[tmp$trialNum == 2] - tmp$freq[tmp$trialNum == 1]
  return(diff)
})

trtDF = sapply(unique(agg_sm$colNum), FUN = function(x){
  tmp = agg_sm[agg_sm$colNum == x, ]
  ttrs = paste(tmp$trt[tmp$trialNum == 1], tmp$trt[tmp$trialNum == 2], sep = "_")
  return(ttrs)
})

buzzdiffs = data.frame(trtDF, diffdf)

ggplot(buzzdiffs, aes(x = trtDF, y= diffdf)) + 
  geom_boxplot() + 
  geom_point()

         
agg2 = aggregate(sl$freq, by=list(colNum = sl$colNum, fullTrt = sl$trt == "full", trt = sl$trt), FUN=mean, na.rm=TRUE)
colnames(agg2)[colnames(agg2) == "x"] = "freq"
agg2$trt = as.character(agg2$trt)

diffdf <- t(as.data.frame(t(sapply(unique(agg2$colNum), FUN = function(x){
  tmp = agg2[agg2$colNum == x, ]
  if(nrow(tmp) <= 1)
    return(NA)
  if (length(unique(tmp$trt)) > 2){
    tmp = tmp[tmp$trt != "full_2", ]
  }
  diff =  tmp$freq[!tmp$fullTrt] - tmp$freq[tmp$fullTrt]
  return(diff)

}))))

length(diffdf)

trtDF = sapply(unique(agg2$colNum), FUN = function(x){
  tmp = agg2[agg2$colNum == x, ]
  if (length(unique(tmp$trt)) > 2){
    tmp = tmp[tmp$trt != "full_2", ]
  }
  ttrs = paste(tmp$trt[tmp$fullTrt], tmp$trt[!tmp$fullTrt], sep = "_")
  return(ttrs)
})

length(trtDF)

buzzdiffs = data.frame(trtDF, diffdf)
tapply(buzzdiffs$diffdf, INDEX = buzzdiffs$trtDF, mean)

ggplot(droplevels(buzzdiffs[buzzdiffs$trtDF != "full_", ]), aes(x = trtDF, y= as.numeric(diffdf))) + 
  geom_boxplot() + 
  geom_point() + 
  labs(y = "diff in frequency -- averages")


m1 = lm(as.numeric(diffdf) ~ trtDF - 1, data = droplevels(buzzdiffs[buzzdiffs$trtDF != "full_", ]))
summary(m1)

sl$trt = relevel(factor(sl$trt), ref = "full")


sl$trt2 = mapvalues(sl$trt, from = c("full", "full_2", "high", "low"), 
                    to = c("full", "full", "high", "low"))


# summary for paper
# fit a varying slope and intercept for colNum (bee ID), and allow the slope of the trialNum
# variable to vary by colNum (beeID)
m2 = lmer(freq ~ trt2 + hive + trialNum + IT_imputed + (1+trialNum|colNum), data = sl, REML = FALSE)

# this model estimates a global intercept
# random effect intercept for colNum (beeID)
# a single global estimate for trialNum
# the effect of trialNum within each level of colNum
# the correlation between intercept of trialNum across levels of colNum

m2_1 = lmer(freq ~ trt2* IT_imputed + hive + trialNum  + (1+trialNum|colNum), data = sl, REML = FALSE)

# compare models with BIC
BIC(m2, m2_1) # m2 (no interaction) is better
anova(m2, m2_1) # Note: this disagrees with BIC

m3 = update(m2, .~. - hive)
BIC(m2, m3) # m3 is better (without hive)
anova(m2, m3) # Note: disagrees with BIC

# refit best model with REML = TRUE
m3 <- update(m3, .~., REML = TRUE)
summary(m3) # summary for paper
plot(m3)
qqnorm(ranef(m3)$colNum[[1]])
qqline(ranef(m3)$colNum[[1]])

sjp.lmer(m3) # plot random effects to find any outliers
sjp.lmer(m3, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects


# post-hoc tests -- pvals for paper
summary(glht(m3, linfct = mcp(trt2 = "Tukey")), test = adjusted("none"))
```


### Generate CI's
```{r bootstrap, cache = TRUE}

# set number of bootstrap samples
nbootSims = 1000

ITmean = mean(tapply(sl$IT_imputed, INDEX = sl$beeCol, FUN = function(x) x[1] ))

# don't need hive, because that's not in the model we chose (above)
pframe <- data.frame(trt2 = levels(droplevels(sl$trt2)), IT_imputed = ITmean,  colNum = 99999, trialNum = 2)
pframe$freq <- 0
pp <- predict(m3, newdata = pframe, re.form=NA, type = 'response') # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(m3, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<-bb2_se[1,]
pframe$bhi<-bb2_se[2,]
pframe$predMean <- pp
pframe <- pframe[, c('trt2', "blo", "bhi", "predMean")]
pframe
```

### Make plots for paper

```{r}
#plot 95% confidence intervals
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
# rename labels for plot
pframe$trt3 = mapvalues(pframe$trt2, from = c("full", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", "High range\n(340 - 390 Hz)", "Low range\n(220 - 330 Hz)"))

pframe
g0 <- ggplot(pframe, aes(x=trt3, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication frequency (Hz)", x = "Frequency range for reward") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)+
  annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 360, label=c("a", "b", "c"),
                color="black") 
g0
ggsave(plot = g0, filename = file.path(figDir, "SonicationFreqPredsAndCI_unadjusted.pdf"), width = 5, height = 4)



g0 <- ggplot(pframe, aes(x=trt3, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication frequency (Hz)", x = "Frequency range for reward") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)+
  annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 360, label=c("a", "b", "b"),
                color="black") 
g0
ggsave(plot = g0, filename = file.path(figDir, "SonicationFreqPredsAndCI_adjustedPvals.pdf"), width = 5, height = 4)



# use the "effects" library to get CI's
# this basically gives the same result as bootstrapping
# it's a good double check
ee <- as.data.frame(Effect(c("trt2"), 
                           fixed.predictors =list(given.values = c(trialNum = 2, 
                                                                   IT_imputed = ITmean)),  
                           m3) )


# compare two methods -- very similar
ee
pframe

g1 <- ggplot(ee, aes(x=trt2, y=fit))+
     geom_point()+ 
     labs(y = "Sonication Frequency (Hz)", x = "Treatment") + 
     geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1)+
     theme_classic() + 
    annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 360, label=c("a", "b", "c"),
                color="black") 
g1

```

### Analysis for amplitude

```{r ampAnalysis}

# refref: here try log amp_acc and REML = FALSE with BIC

# interaction model
# note log-transformation to make model fit assumptions better
maa0 = lmer(log(amp_acc) ~ trt2* IT_imputed + hive + trialNum  + (1+trialNum|colNum), data = sl, REML = FALSE)

# main effect model
maa1 = lmer(log(amp_acc) ~ trt2 + hive + trialNum + IT_imputed + (1+trialNum|colNum), data = sl, REML = FALSE)
BIC(maa0, maa1) # use no interaction (keep maa1)
anova(maa0, maa1) # note that this agrees with likelihood ratio test


maa2 = update(maa1, .~. - trt2)
BIC(maa1, maa2) # keep treatment (maa1)
anova(maa1, maa2) # agrees with LRT; p-val for paper, if needed

maa3 <- update(maa1, .~. - hive)
BIC(maa1, maa3) # remove hive (maa3)
anova(maa1, maa3) # agrees with LRT


# refit with REML = TRUE for paper
maa3 <- update(maa3, .~., REML = TRUE)
summary(maa3)


# diagnostics
plot(maa3)
qqnorm(ranef(maa1)$colNum[[1]])
qqline(ranef(maa1)$colNum[[1]])


sjp.lmer(maa3) # plot random effects to find any outliers
sjp.lmer(maa3, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects


# post-hoc tests -- pvals for paper
summary(glht(maa3, linfct = mcp(trt2 = "Tukey")), test = adjusted("none"))


# post-hoc tests with bonf adjustment
summary(glht(maa3, linfct = mcp(trt2 = "Tukey")), test = adjusted("bonf"))
```


### Generate CI's for amplitude

```{r bootstrapAmp, cache = TRUE}
# set number of bootstrap samples
nbootSims2 = 1000

# don't need to include hive, but will set trialNum to 2
pframe <- data.frame(trt2 = levels(droplevels(sl$trt2)), 
                     IT_imputed = ITmean,  colNum = 99999, trialNum = 2)
pframe$amp_acc <- 0
# exponentiate to get back to original scale
pp <- exp(predict(maa3, newdata = pframe, re.form=NA, type = 'response')) # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(maa3, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims2)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))

#exponentiate to get back to orignal scale
pframe$blo<-exp(bb2_se[1,])
pframe$bhi<-exp(bb2_se[2,])
pframe$predMean <- pp
pframe <- pframe[, c('trt2', "blo", "bhi", "predMean")]
pframe
```


### Make plots for amplitude for paper

```{r amplitudePlots}
#plot 95% confidence intervals
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
pframe$trt3 = mapvalues(pframe$trt2, from = c("full", "high", "low"), 
                        to = c("Full range\n(220 - 450 Hz)", "High range\n(340 - 390 Hz)", "Low range\n(220 - 330 Hz)"))

pframe
ga0 <- ggplot(pframe, aes(x=trt3, y=predMean))+
     geom_point()+ 
     labs( y = expression ("Sonication amplitude "(m~s^{-2})), x = "Frequency range for reward") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)+
     theme(axis.text.x = element_text(angle = 45, hjust = 1), 
           legend.position = 'none') +
     theme_classic() + 
  annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 56, label=c("a", "b", "c"),
                color="black") 
ga0
ggsave(plot = ga0, filename = file.path(figDir, "SonicationAmpPredsAndCI_unadjusted.pdf"), width = 5, height = 4)


# make plot using bonferroni adjustments

ga0 <- ggplot(pframe, aes(x=trt3, y=predMean))+
     geom_point()+ 
     labs( y = expression ("Sonication amplitude "(m~s^{-2})), x = "Frequency range for reward") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)+
     theme(axis.text.x = element_text(angle = 45, hjust = 1), 
           legend.position = 'none') +
     theme_classic() + 
  annotate(geom="text", x=c(1,2,3), y=c(0, 0, 0) + 56, label=c("a", "b", "a"),
                color="black") 
ga0
ggsave(plot = ga0, filename = file.path(figDir, "SonicationAmpPredsAndCI_BonfAdjusted.pdf"), width = 5, height = 4)

```


# Visualize aggregated data for amplitude

```{r}

aggdata <- aggregate(sl$amp_acc, by=list(colNum = sl$colNum,trialNum =sl$trialNum, trt = sl$trt), FUN=mean, na.rm=TRUE)
colnames(aggdata)[colnames(aggdata) == "x"] = "amp"

aggdata_sd <- aggregate(sl$amp_acc, by=list(colNum = sl$colNum,trialNum =sl$trialNum, trt = sl$trt), FUN=sd, na.rm=TRUE)
colnames(aggdata_sd)[colnames(aggdata_sd) == "x"] = "amp_sd"

aggdata = merge(aggdata, aggdata_sd)
aggdata = aggdata[order(aggdata$colNum, aggdata$trialNum, decreasing = FALSE), ]
agg_sm = aggdata[aggdata$trialNum <= 2, ]
rownames(agg_sm) = 1:nrow(agg_sm)
agg_sm



agg_sm[duplicated(data.frame(agg_sm$colNum, agg_sm$trialNum)), ]

ggplot(agg_sm, aes(x = trt, y = amp, fill = trialNum > 1)) +
  geom_boxplot(alpha = 0.2) + 
  geom_point(aes(color = trialNum>1)) + 
  geom_line(aes(group = colNum))


diffdf <- sapply(unique(agg_sm$colNum), FUN = function(x){
  tmp = agg_sm[agg_sm$colNum == x, ]
  if(nrow(tmp) <= 1)
    diff = NA
  else
    diff = tmp$amp[tmp$trialNum == 2] - tmp$amp[tmp$trialNum == 1]
  return(diff)
})

trtDF = sapply(unique(agg_sm$colNum), FUN = function(x){
  tmp = agg_sm[agg_sm$colNum == x, ]
  ttrs = paste(tmp$trt[tmp$trialNum == 1], tmp$trt[tmp$trialNum == 2], sep = "_")
  return(ttrs)
})

buzzdiffs = data.frame(trtDF, diffdf)

ggplot(buzzdiffs, aes(x = trtDF, y= diffdf)) + 
  geom_boxplot() + 
  geom_point()+ 
  labs(y = "amplitude difference m/s/s")

         
agg2 = aggregate(sl$amp_acc, by=list(colNum = sl$colNum, fullTrt = sl$trt == "full", trt = sl$trt), FUN=mean, na.rm=TRUE)
colnames(agg2)[colnames(agg2) == "x"] = "amp"
agg2$trt = as.character(agg2$trt)

diffdf <- t(as.data.frame(t(sapply(unique(agg2$colNum), FUN = function(x){
  tmp = agg2[agg2$colNum == x, ]
  if(nrow(tmp) <= 1)
    return(NA)
  if (length(unique(tmp$trt)) > 2){
    tmp = tmp[tmp$trt != "full_2", ]
  }
  diff =  tmp$amp[!tmp$fullTrt] - tmp$amp[tmp$fullTrt]
  return(diff)

}))))

length(diffdf)

trtDF = sapply(unique(agg2$colNum), FUN = function(x){
  tmp = agg2[agg2$colNum == x, ]
  if (length(unique(tmp$trt)) > 2){
    tmp = tmp[tmp$trt != "full_2", ]
  }
  ttrs = paste(tmp$trt[tmp$fullTrt], tmp$trt[!tmp$fullTrt], sep = "_")
  return(ttrs)
})

length(trtDF)

buzzdiffs = data.frame(trtDF, diffdf)
tapply(buzzdiffs$diffdf, INDEX = buzzdiffs$trtDF, mean)

ggplot(droplevels(buzzdiffs[buzzdiffs$trtDF != "full_", ]), aes(x = trtDF, y= as.numeric(diffdf))) + 
  geom_boxplot() + 
  geom_point()+ 
  labs(y = "amplitude difference m/s/s")
```


## Plot amplitude vs. frequency for initial trial, accounting for size

```{r}
ggplot(sl[sl$trialNum == 1, ], aes(x = freq, y = amp_acc)) + 
  geom_point(position = position_jitter(width = 3), alpha = 0.2) + 
  theme_classic() + 
  geom_smooth(aes(group = trt2)) + 
  labs( y = expression ("Sonication amplitude "(m~s^{-2})), x = "Frequency (Hz)")

centerFreq = scale(sl[sl$trialNum == 1, "freq"])

m1 <- lmer(log(amp_acc)~centerFreq + I(centerFreq^2) + I(centerFreq^3) + IT_imputed + (1|beeCol), data = sl[sl$trialNum == 1, ])
summary(m1)

plot(m1)

# predict -- using mean IT span
ppdf <- data.frame(centerFreq = sort(unique(centerFreq)), IT_imputed = ITmean, beeCol = 99999, acc_amp = 0)
# exponentiate to get back to original scale
ppdf$predAmp = exp(predict(m1, newdata = ppdf, type = "response", re.form = NA))



cf_Unscaled = ppdf$centerFreq * attr(centerFreq, 'scaled:scale') + attr(centerFreq, 'scaled:center')
ggplot(ppdf, aes(x = cf_Unscaled, y = predAmp))+ 
  geom_line() + 
  labs(x = "Sonication Frequency (Hz)", y = expression ("Sonication amplitude "(m~s^{-2})) ) + 
  geom_point(data = sl[sl$trialNum == 1, ], 
             aes(x = freq, y = amp_acc), 
             alpha = 0.2, position = position_jitter(width =2), pch =16, size = .5) 
ggsave(filename = file.path(figDir, "FreqVsAmp_1stTrial_rawData.pdf"), width = 4, height = 3)

g22 <- ggplot(ppdf, aes(x = cf_Unscaled, y = predAmp))+ 
  geom_line() + 
  labs(x = "Sonication Frequency (Hz)", y = expression ("Predicted sonication amplitude "(m~s^{-2})) ) 
g22
ggsave(filename = file.path(figDir, "FreqVsAmp_1stTrial.pdf"), width = 4, height = 3)
  
```

```{r}
sessionInfo()
```

