---
output: html_document
editor_options: 
  chunk_output_type: console
---

<pre>
# Heavy vs. light flowers
# Callin Switzer
# 12 October 2016
# update 8 Dec 2017
# 
# Compares the bees' frequency when they are buzzing on 
# heavy (metal added) vs. light flowers
# Note: all pores were glued shut
</pre>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/cswitzer/Dropbox/SonicationBehavior")
```



```{r}
#install packages
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("ggplot2", "reshape2", 'lme4', 'sjPlot', "multcomp", "plyr", "effects", "gamm4", "viridis")
ipak(packages)

# set ggplot theme
theme_set(theme_classic() + theme(axis.text=element_text(colour="black")))

# define data and figure directories
dataDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehData"
figDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehFigs"


# print system info
print(paste("last run ", Sys.time()))
print(R.version)
```


### Read in data and do some visualizations

```{r}
new_df = read.csv(file.path(dataDir, "02_HeavyLight_cleaned.csv"))
new_df$hive = as.factor(new_df$hive)
new_df$treatment <- mapvalues(new_df$treatment, from = c("sham", "weighted"), to = c("Sham", "Weighted"))
head(new_df)

# calculate average number of each treatment
ll = list(tapply(new_df$treatment, new_df$beeID, table))
dd = data.frame(unlist(ll))
dd$beeTrt = row.names(dd)
dd$beeID = sapply(dd$beeTrt, FUN = function(x) strsplit(x, "\\.")[[1]][1])
dd$trt = sapply(dd$beeTrt, FUN = function(x) strsplit(x, "\\.")[[1]][2])

# shows average number of buzzes per treatment
tapply(dd$unlist.ll., INDEX = dd$trt, mean)
tapply(dd$unlist.ll., INDEX = dd$trt, sd) # sd for number of buzzes

# show histograms of frequencies
hist(new_df$freq, breaks = seq(215, 455, by = 10))

# Calculate average for each bee and for each treatment
frqMeans <- as.data.frame(tapply(X = new_df$freq, INDEX = list(new_df$beeID, new_df$treatment), mean))
frqMeans$beeID <- row.names(frqMeans)
frqMeans


# convert to long format for ggplot-ing
frqLong <- melt(frqMeans, id.vars = "beeID", measure.vars =  c("Sham", "Weighted"), 
     variable.name = "trt", value.name = "frq")
nrow(frqLong)

ggplot(frqLong, aes(x = trt, y = frq)) + 
     geom_boxplot() + 
    geom_line(aes(group = beeID))+
     geom_point()


ggplot(frqLong, aes(x = trt, y = frq)) + 
     geom_boxplot() + 
     labs(y = "Buzz Frequency (Hz)", x = "Flower treatment")
#ggsave(file.path(figDir, "heavyLight_frq.pdf"), width = 5, height = 4)

# print some descriptive statistics
nrow(new_df)
unique(new_df$hive)



# Calculate average amplitude for each bee and for each treatment
ampMeans <- as.data.frame(tapply(X = new_df$amp_acc, INDEX = list(new_df$beeID, new_df$treatment), mean))

ampMeans$beeID <- row.names(ampMeans)
ampMeans


ampLong <- melt(ampMeans, id.vars = "beeID", measure.vars =  c("Sham", "Weighted"), 
                variable.name = "trt", value.name = "amp")

ggplot(ampLong, aes(x = trt, y = amp)) + 
     geom_boxplot() + 
     geom_point()  + 
  labs(y = expression ("Sonication amplitude "(m~s^{-2})), x = "Flower treatment")

ggplot(ampLong, aes(x = trt, y = amp)) + 
     geom_boxplot() + 
     geom_point() +
  geom_line(aes(group = beeID))


ggplot(ampLong, aes(x = trt, y = amp)) + 
     geom_boxplot() + 
     labs(y = expression ("Sonication amplitude "(m~s^{-2})), x = "Flower treatment")
#ggsave(file.path(figDir, 'heavyLightAmp.pdf'), width = 5, heigh = 4)

nrow(ampLong)
```


### Modeling frequency with GAMM

```{r}
rr = new_df$treatment[1]
df1 = new_df[1, ]
for(ii in 1:(nrow(new_df) -1)){
  if(rr == new_df$treatment[ii + 1]) next
  else{
    df1  = rbind(df1, new_df[ii+1, ])
    rr = new_df$treatment[ii + 1]
  }
}

df1 <- df1[df1$rewNum != 1, ]
df1$trtSwitch = df1$rewNum
colnames(df1)
df2 <- df1[, c("beeID", "hive", "trtSwitch", "IT")]
df2

sum(duplicated(df2$beeID))

newdf2 <- merge(new_df, df2)
head(newdf2)


newdf2$visitNum_centered <- newdf2$rewNum - newdf2$trtSwitch

# get treatment
trts <- (sapply(unique(newdf2$beeID), FUN = function(x){
  startt = newdf2[newdf2$beeID == x & newdf2$rewNum == 1, "treatment" ]
  return(as.character(startt))
}))

df4 <- data.frame(beeID = unique(newdf2$beeID), trts = trts)

newdf3 <- merge(newdf2, df4)

newdf3$trt2 <- as.factor(paste(newdf3$trts, "first"))

head(newdf3)


# GAMM
# start with gamm so I can show change by visit number
g01 = gamm4(freq ~ s(visitNum_centered, by = trt2) + IT + hive, random =  ~(1|beeID), data = newdf3)
par(mfrow = c(2,3))
aab <- plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g01$gam) # Summary for paper 
summary(g01$mer)
dev.off()

# looks like GAMM is not necessary -- predictions are linear
```


### Modeling frequency with lmer

Use BIC to select model (decide what interactions and covariates to use)

Make sure that REML = FALSE when comparing BIC values

Start with the biggest model of interest, and then see what predictors can be removed

```{r modeling}
# fit lmer that is similar to gamm above
# interaction that may be important, based on domain knowledge
m0 = lmer(freq ~ I(scale(visitNum_centered)) * trt2 + IT + hive  + I(scale(visitNum_centered^2)) + treatment + (1|beeID), data = newdf3, REML = FALSE)
summary(m0)

# no interaction
m1 = update(m0, .~. - I(scale(visitNum_centered)) : trt2)
BIC(m0, m1) # stay with m1 -- no interaction
anova(m0, m1) # agrees with BIC

summary(m1)

m1.1 <- update(m1, .~. - I(scale(visitNum_centered^2)))
anova(m1.1, m1) # keep 1.1
summary(m1.1)

m2 <- update(m1.1, .~. - IT)
anova(m1.1, m2) # keep m2 (based on BIC)

summary(m2)

m3 <- update(m2, .~. - hive)
anova(m2, m3) # keep m3

summary(m3)

m3.1 <- update(m3, .~. - treatment)
anova(m3, m3.1) # keep m3.1

m4 <- update(m3.1, .~. - trt2)
summary(m4)
anova(m3.1, m4) #keep m4


m5 <- update(m4, .~. -I(scale(visitNum_centered)))
anova(m4, m5) # keep m4

m4 <- update(m4, .~., REML = TRUE)
summary(m4) # final model for paper


# diagnostics -- use REML = TRUE
m1 <- update(m4, .~., REML =TRUE)
summary(m1) # summary for paper
plot(m1)
qqnorm(ranef(m1)$beeID[[1]])
qqline(ranef(m1)$beeID[[1]])

sjp.lmer(m1, type = "re", sort = TRUE) # plot random effects to find any outliers
sjp.lmer(m1, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects
```


### Bootstrap CI's for figure for paper

```{r bootstrap, cache = TRUE}
# set number of bootstrap replicates for models
nbootSims = 1000

table(new_df$hive) # more trials from hive 3

# using hive 3, since it's the one with the most data
# however, hive doesn't affect model anyway

# calculate an average IT for prediction
ITmean = mean(tapply(new_df$IT, INDEX = new_df$beeID, FUN = function(x) x[1] ))

pframe <- expand.grid(trt2 = levels(droplevels(newdf3$trt2)),
                     IT = ITmean,
                     visitNum_centered = -50:50, 
                     hive = factor(3, levels = levels(newdf3$hive)),  
                     beeID = 99999, 
                     treatment = levels(newdf3$treatment))
pframe$freq <- 0
pframe <- pframe[(pframe$trt2 == "Sham first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Sham first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered > 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered > 0) 
                    , ]


m1 <- update(m1, .~. + treatment)
pp <- predict(m1, newdata = pframe, re.form=NA, type = 'response') # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(m1, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<-bb2_se[1,]
pframe$bhi<-bb2_se[2,]
pframe$predMean <- pp
pframe
```

### Make frequency plots for paper

```{r figure-making}

g0 <- ggplot(pframe, aes(x=visitNum_centered, y=predMean))+
     geom_line(aes(color = treatment))+ 
     labs(y = "Frequency (Hz)", x = "Sonication number\n (0 is when treatment switched)") + 
     geom_ribbon(aes(x = visitNum_centered, ymin = blo, ymax = bhi, fill= treatment), alpha = 0.2) + 
  facet_wrap(~trt2) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  theme(legend.position = c(0.5, 0.90), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") + 
  ylim(c(330, 370))
g0
ggsave(filename = file.path(figDir, "WeightedSham_Freq_Timeseries.pdf"), width =5, height = 3.6)





# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
g0 <- ggplot(pframe, aes(x=visitNum_centered, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication frequency (Hz)", x = "buzzNumber") + 
     geom_ribbon(aes(x = visitNum_centered, ymin = blo, ymax = bhi), alpha = 0.2) + 
  facet_wrap(~trt2)
g0

# plot effect to see if it agrees
plot(Effect(c("visitNum_centered"), m1))


# make figure that doesn't show visitNum
# set number of bootstrap replicates for models
nbootSims = 100

m1 = update(m1, .~. - I(scale(visitNum_centered)) : treatment, REML = TRUE)

table(new_df$hive) # more trials from hive 3

# using hive 3, since it's the one with the most data
# however, hive doesn't affect model anyway

# calculate an average IT for prediction
ITmean = mean(tapply(new_df$IT, INDEX = new_df$beeID, FUN = function(x) x[1] ))

pframe <- expand.grid(trt2 = levels(droplevels(newdf3$trt2)),
                     IT = ITmean,
                     visitNum_centered = c(-1,0,1), 
                     hive = factor(3, levels = levels(newdf3$hive)),  
                     beeID = 99999, 
                     treatment = levels(newdf3$treatment))
pframe$freq <- 0

pframe2 <- pframe[(pframe$trt2 == "Sham first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Sham first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered > 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered > 0) 
                    , ]


pp <- predict(m1, newdata = pframe2, re.form=NA, type = 'response') # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(m0, FUN=function(x) predict(x, pframe2, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe2$blo<-bb2_se[1,]
pframe2$bhi<-bb2_se[2,]
pframe2$predMean <- pp
pframe2




g0 <- ggplot(pframe, aes(x=visitNum_centered, y=predMean))+
     geom_point()+ 
     labs(y = "amplitude", x = "buzzNumber") + 
     geom_errorbar(aes( ymin = blo, ymax = bhi), width = 0.1) + 
  facet_wrap(~trt2)
g0


```



### Modeling amplitude with gamm

```{r modeling-amp}
g01 = gamm4(log(amp_acc) ~ s(visitNum_centered, by = trt2), random =  ~(1|beeID), data = newdf3)
par(mfrow = c(2,3))
aab <- plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g01$gam) # Summary for paper 
summary(g01$mer)
dev.off()
```



### Modeling amplitude with lmer

```{r modeling-amp}
# log transform for acceleration helps model fit better
# start with large model (including interaction)
ma0 = lmer(log(amp_acc) ~ visitNum_centered*trt2 +  hive + treatment + IT+ (1|beeID), data = newdf3, REML = FALSE)
summary(ma0)

ma1 = update(ma0, .~. - visitNum_centered:trt2)

BIC(ma0, ma1) # interaction, ma1, is better

# pval for interaction (in case paper needs it)
anova(ma0, ma1)
summary(ma1)

ma2 <- update(ma1, .~. - hive)
BIC(ma1, ma2) # hive not important
summary(ma2)


ma3 <- update(ma2, .~. - trt2)
anova(ma2, ma3) # keep ma3

summary(ma3)

ma3.1 <- update(ma3, .~. - IT)
anova(ma3, ma3.1) # keep 3.1


ma4 <- update(ma3.1, .~. + visitNum_centered:treatment)
anova(ma3.1, ma4) #keep ma3

ma3 <- update(ma3.1, .~., REML = TRUE)

# diagnostics
plot(ma3)
qqnorm(ranef(ma3)$beeID[[1]])
qqline(ranef(ma3)$beeID[[1]])


sjp.lmer(ma3) # plot random effects to find any outliers
sjp.lmer(ma3, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects
```


### Bootstrap CI's for amplitude for figure for paper

```{r bootstrap_amp, cache = TRUE}

# set number of bootstrap replicates for models
nbootSims2 = 1000

# using hive 3, since it's the one with the most data
pframe <- expand.grid(trt2 = levels(droplevels(newdf3$trt2)),
                     IT = ITmean,
                     visitNum_centered = -50:50, 
                     hive = factor(3, levels = levels(newdf3$hive)),  
                     beeID = 99999, 
                     treatment = levels(droplevels(newdf3$treatment)))
pframe$amp_acc <- 0


pframe2 <- pframe[(pframe$trt2 == "Sham first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Sham first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered > 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered > 0) 
                    , ]
# exponentiate to put on original scale
pp <- exp(predict(ma3, newdata = pframe2, re.form=NA, type = 'response')) # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(ma3, FUN=function(x) predict(x, pframe2, re.form=NA, type = 'response'), nsim = nbootSims2)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe2$blo<- exp(bb2_se[1,]) # exponentiate to put on original scale
pframe2$bhi<- exp(bb2_se[2,])
pframe2$predMean <- pp
pframe2
```



### Plot amplitude lmer with trial number
```{r}
g0 <- ggplot(pframe2, aes(x=visitNum_centered, y=predMean))+
     geom_line(aes(color = treatment))+ 
     labs(y = expression ("Sonication amplitude "(m~s^{-2})), x = "Sonication number\n (0 is when treatment switched)") + 
     geom_ribbon(aes(x = visitNum_centered, ymin = blo, ymax = bhi, fill= treatment), alpha = 0.2) + 
  facet_wrap(~trt2) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  theme(legend.position = c(0.5, 0.80), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="vertical")
g0


ggsave(filename = file.path(figDir, "WeightedSham_Timeseries.pdf"), width =5, height = 3.6)
# I like this plot

```



### Make amplitude plots for paper

```{r figure-making-amp}
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
# Holding IT = meanIT
print(ITmean)
ga0 <- ggplot(pframe, aes(x=treatment, y=predMean))+
     geom_point()+ 
     labs(y = expression ("Sonication amplitude "(m~s^{-2})), x = "Flower treatment") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1) +
  annotate(geom="text", x=c(1,2), y=c(0, 0) + 40, label=c("a", "b"),
                color="black") 
ga0
ggsave(plot = ga0, filename = file.path(figDir, "HeavyLight_PredsAndCI_amp.pdf"), width =4, height = 3)
```

```{r}
sessionInfo()
````

