---
output: html_document
editor_options: 
  chunk_output_type: console
---

<pre>
# Heavy vs. light flowers
# Callin Switzer
# 12 October 2016
# update 8 Dec 2017
# update 9 Aug 2018 -- make figures svg

# Compares the bees' frequency when they are buzzing on 
# heavy (metal added) vs. light flowers
# Note: all pores were glued shut
</pre>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# define data and figure directories
windows = ifelse(grepl(pattern = "wind", ignore.case = TRUE, 
                       x = Sys.info()[1]), TRUE, FALSE)
  
{if(windows){
  rd = "D:/Dropbox/SonicationBehavior"
}
else{
  rd = "/Users/cswitzer/Dropbox/SonicationBehavior"
}}

knitr::opts_knit$set(root.dir =rd )
```


```{r}
#install packages
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("ggplot2", "reshape2", 'lme4', 'sjPlot', "multcomp", "plyr", "effects", "gamm4", "viridis", 
              "cowplot")
ipak(packages)

# set ggplot theme
theme_set(theme_classic() + theme(axis.text=element_text(colour="black")))

# define data and figure directories
{if(windows){
  # windows
  dataDir <- "D:/Dropbox/SonicationBehavior/SonBehData"
  figDir <- "D:/Dropbox/SonicationBehavior/SonBehFigs"
  }
else{
  # mac
  dataDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehData"
  figDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehFigs"
}}

# check if the directory exists
file.info(dataDir)$isdir


# print system info
print(paste("last run ", Sys.time()))
print(R.version)
```

### Read in data and do some visualizations

```{r}
#new_df = read.csv(file.path(dataDir, "02_HeavyLight_cleaned.csv"))
new_df = read.csv(file.path(dataDir, "02_1_HeavyLight_cleaned_posAdded.csv"))

new_df$amp = new_df$MinMax_amp_Volts
new_df$amp_acc = new_df$MinMax_amp_acc_m.s.s

# divide amplitude by two
new_df$amp = new_df$amp / 2
new_df$amp_acc = new_df$amp_acc / 2

new_df$hive = as.factor(new_df$hive)
new_df$treatment <- mapvalues(new_df$treatment, from = c("sham", "weighted"), to = c("Sham", "Weighted"))
head(new_df)

# calculate average number of each treatment
ll = list(tapply(new_df$treatment, new_df$beeID, table))
dd = data.frame(unlist(ll))
dd$beeTrt = row.names(dd)
dd$beeID = sapply(dd$beeTrt, FUN = function(x) strsplit(x, "\\.")[[1]][1])
dd$trt = sapply(dd$beeTrt, FUN = function(x) strsplit(x, "\\.")[[1]][2])

# shows average number of buzzes per treatment
tapply(dd$unlist.ll., INDEX = dd$trt, mean)
tapply(dd$unlist.ll., INDEX = dd$trt, sd) # sd for number of buzzes

# show histograms of frequencies
hist(new_df$freq, breaks = seq(215, 455, by = 10))

# Calculate average for each bee and for each treatment
frqMeans <- as.data.frame(tapply(X = new_df$freq, INDEX = list(new_df$beeID, new_df$treatment), mean))
frqMeans$beeID <- row.names(frqMeans)
frqMeans


# convert to long format for ggplot-ing
frqLong <- melt(frqMeans, id.vars = "beeID", measure.vars =  c("Sham", "Weighted"), 
     variable.name = "trt", value.name = "frq")
nrow(frqLong)

ggplot(frqLong, aes(x = trt, y = frq)) + 
     geom_boxplot() + 
    geom_line(aes(group = beeID))+
     geom_point()


ggplot(frqLong, aes(x = trt, y = frq)) + 
     geom_boxplot() + 
     labs(y = "Buzz Frequency (Hz)", x = "Flower treatment")
#ggsave(file.path(figDir, "heavyLight_frq.pdf"), width = 5, height = 4)

# print some descriptive statistics
nrow(new_df)
unique(new_df$hive)



# Calculate average amplitude for each bee and for each treatment
ampMeans <- as.data.frame(tapply(X = new_df$amp_acc, INDEX = list(new_df$beeID, new_df$treatment), mean))

ampMeans$beeID <- row.names(ampMeans)
ampMeans


ampLong <- melt(ampMeans, id.vars = "beeID", measure.vars =  c("Sham", "Weighted"), 
                variable.name = "trt", value.name = "amp")

ggplot(ampLong, aes(x = trt, y = amp)) + 
     geom_boxplot() + 
     geom_point()  + 
  labs(y = expression ("Sonication acceleration "(m~s^{-2})), x = "Flower treatment")

ggplot(ampLong, aes(x = trt, y = amp)) + 
     geom_boxplot() + 
     geom_point() +
  geom_line(aes(group = beeID))


ggplot(ampLong, aes(x = trt, y = amp)) + 
     geom_boxplot() + 
     labs(y = expression ("Sonication acceleration "(m~s^{-2})), x = "Flower treatment")
#ggsave(file.path(figDir, 'heavyLightAmp.pdf'), width = 5, heigh = 4)

nrow(ampLong)
```


### Modeling frequency with GAMM

Use BIC to select model (decide what interactions and covariates to use)

Make sure that REML = FALSE when comparing BIC values

Start with the biggest model of interest, and then see what predictors can be removed

```{r}
rr = new_df$treatment[1]
df1 = new_df[1, ]
for(ii in 1:(nrow(new_df) -1)){
  if(rr == new_df$treatment[ii + 1]) next
  else{
    df1  = rbind(df1, new_df[ii+1, ])
    rr = new_df$treatment[ii + 1]
  }
}

df1 <- df1[df1$rewNum != 1, ]
df1$trtSwitch = df1$rewNum
colnames(df1)
df2 <- df1[, c("beeID", "hive", "trtSwitch", "IT")]
df2

sum(duplicated(df2$beeID))

newdf2 <- merge(new_df, df2)
head(newdf2)


newdf2$visitNum_centered <- newdf2$rewNum - newdf2$trtSwitch

# get treatment
trts <- (sapply(unique(newdf2$beeID), FUN = function(x){
  startt = newdf2[newdf2$beeID == x & newdf2$rewNum == 1, "treatment" ]
  return(as.character(startt))
}))

df4 <- data.frame(beeID = unique(newdf2$beeID), trts = trts)

newdf3 <- merge(newdf2, df4)

newdf3$trt2 <- as.factor(paste(newdf3$trts, "first"))

head(newdf3)


# GAMM
# start with gamm so I can show change by visit number
g01 = gamm4(freq ~ s(visitNum_centered, by = trt2) + IT + hive + treatment + trt2, random =  ~(1|beeID), data = newdf3)
par(mfrow = c(2,3))
aab <- plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g01$gam) # Summary for paper 
summary(g01$mer)
dev.off()


g02 = gamm4(freq ~ s(visitNum_centered, by = trt2) +  hive + treatment + trt2, random =  ~(1|beeID), data = newdf3)
anova(g01$mer, g02$mer) #g02 better
summary(g02$mer)


g03 =  gamm4(freq ~ s(visitNum_centered, by = trt2) +  hive + trt2, random =  ~(1|beeID), data = newdf3)
anova(g02$mer, g03$mer) #g03 better
summary(g03$mer)


g04 = gamm4(freq ~ s(visitNum_centered, by = trt2) +trt2, random =  ~(1|beeID), data = newdf3)
anova(g03$mer, g04$mer) # g04 better
summary(g04$mer)

g05 = gamm4(freq ~ s(visitNum_centered) +trt2, random =  ~(1|beeID), data = newdf3)
anova(g04$mer, g05$mer) # g05 slightly better (according to BIC)

summary(g05$gam)


g06 = gamm4(freq ~ s(visitNum_centered), random =  ~(1|beeID), data = newdf3)
anova(g05$mer, g06$mer) # g05 slightly better (according to BIC)

summary(g06$mer) # g06 Better

g07 = gamm4(freq ~ 1, random =  ~(1|beeID), data = newdf3)
anova(g06$mer, g07$mer) # g07 better

# looks like GAMM is not necessary -- no predictors significantly affect frequency
```


### Modeling frequency with lmer

#### Do a similar thing to GAMM above (but make it more interpretable)

```{r modeling1}
# fit lmer that is similar to gamm above
# interaction that may be important, based on domain knowledge
m0 = lmer(freq ~ I(scale(visitNum_centered)) * trt2 + I(scale(IT)) * treatment + hive  + I(scale(visitNum_centered^2)) + I(scale(IT)^2) +   (1|beeID), data = newdf3, REML = FALSE)

summary(m0)






# no interaction
m1 = update(m0, .~. - I(scale(visitNum_centered^2)))
BIC(m0, m1) # stay with m1 -- no interaction
anova(m0, m1) # agrees with BIC

summary(m1)

m1.1 <- update(m1, .~. - I(scale(visitNum_centered)):trt2)
anova(m1.1, m1) # keep 1.1
summary(m1.1)

m2 <- update(m1.1, .~. - hive)
anova(m1.1, m2) # keep m2 (based on BIC)

summary(m2)

m3 <- update(m2, .~. - I(scale(IT)^2))
anova(m2, m3) # keep m3

summary(m3)

m3.1 <- update(m3, .~. - trt2)
anova(m3, m3.1) # keep m3.1
summary(m3.1)

m4 <- update(m3.1, .~. - treatment)

anova(m3.1, m4) #keep m4
summary(m4)


m5 <- update(m4, .~. -I(scale(IT)):treatment)
anova(m4, m5) # keep m5 -- baesd on BIC (agrees with GAMM analysis above)

summary(m5)


m6 <- update(m5, .~. - I(scale(IT)))
anova(m5, m6)

summary(m6) 

m7 <- update(m6, .~. - I(scale(visitNum_centered)))
anova(m6, m7) # keep m7

summary(m7)

# get p-value for treament  for paper

m_trt <- update(m7, .~. + treatment)
anova(m7, m_trt) # p-val for paper


m_IT <- update(m7, .~. + IT)
anova(m7, m_IT) # p-val for IT span


# final freq mod for paper
finmod <- update(m7, .~., REML = TRUE)
summary(finmod) # final model for paper -- same as gamm


summary(finmod) # summary for paper -- no predictors predict frequency (when accounting for beeID)
plot(finmod)
qqnorm(ranef(finmod)$beeID[[1]])
qqline(ranef(finmod)$beeID[[1]])

sjp.lmer(finmod, type = "re", sort = TRUE) # plot random effects to find any outliers
# sjp.lmer(finmod, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects
```




### Bootstrap CI's for figure for paper

```{r bootstrap1, cache = TRUE}
# set number of bootstrap replicates for models
nbootSims = 10000

table(new_df$hive) # more trials from hive 3

# using hive 3, since it's the one with the most data
# however, hive doesn't affect model anyway

# calculate an average IT for prediction
ITmean = mean(tapply(new_df$IT, INDEX = new_df$beeID, FUN = function(x) x[1] ))

pframe <- expand.grid(trt2 = levels(droplevels(newdf3$trt2)),
                     IT = ITmean,
                     visitNum_centered = -50:50, 
                     hive = factor(3, levels = levels(newdf3$hive)),  
                     beeID = 99999, 
                     treatment = levels(newdf3$treatment))
pframe$freq <- 0
pframe <- pframe[(pframe$trt2 == "Sham first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Sham first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered > 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered > 0) 
                    , ]

# make a model to show the effect of treatment in a plot
m1 <- update(finmod, .~. + treatment)
pp <- predict(m1, newdata = pframe, re.form=NA, type = 'response') # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(m1, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<-bb2_se[1,]
pframe$bhi<-bb2_se[2,]
pframe$predMean <- pp
pframe

### Make frequency plots for paper


# predictions are from this model
summary(m1)
anova(m1, update(m1, .~. - treatment)) #p-val and BIC for paper

pframe$trt3 <- mapvalues(pframe$trt2, from = c("Sham first", "Weighted first"), 
                          to = c("Sham -> Weighted", "Weighted -> Sham"))


g0 <- ggplot(pframe, aes(x=visitNum_centered, y=predMean))+
     geom_line(aes(color = treatment))+ 
     labs(y = "Sonication frequency (Hz)", x = "Sonication number\n (0 is when treatment switched)") + 
     geom_ribbon(aes(x = visitNum_centered, ymin = blo, ymax = bhi, fill= treatment), alpha = 0.2) + 
  facet_wrap(~trt3) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  theme(legend.position = c(0.5, 0.90), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") + 
  ylim(c(330, 370))
g0
ggsave(filename = file.path(figDir, "WeightedSham_Freq_Timeseries.pdf"), width =5, height = 3.6)


g1 <- g0 + geom_point(data = newdf3[newdf3$visitNum_centered < 50,],
                aes(x = visitNum_centered, y = freq), position = position_jitter(height = 5), size = 0.1, pch = 16, alpha = 0.4) + 
  ylim(c(220, 470))

g1 # note that the distribution of frequency is a bit bimodal -- this could be 
# a problem, so we ran the analysis with a freq cutoff at 270
# we found no significant differences in the models either way.
ggsave(filename = file.path(figDir, "WeightedSham_Freq_Timeseries_RawData.pdf"), width =5, height = 3.6)


g1 <- g0 + geom_point(data = newdf3[newdf3$visitNum_centered < 50,],
                aes(x = visitNum_centered, y = freq), position = position_jitter(height = 5), size = 0.1, pch = 16, alpha = 0.4) + 
  ylim(c(270, 480))

g1
ggsave(filename = file.path(figDir, "WeightedSham_Freq_Timeseries_RawData_trimmed.pdf"), width =5, height = 3.6)



# mean and 95% CI (without visit number)

g0_freq <- ggplot(pframe[pframe$visitNum_centered %in% c(-25, 25), ], aes(x=treatment, y=predMean))+
    geom_point(aes(color = treatment), alpha = 1) + 
     labs(y = "Sonication frequency (Hz)", x = "Flower treatment") + 
     geom_errorbar(aes(x = treatment, ymin = blo, ymax = bhi, color= treatment), alpha = 1, width = 0.1) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, begin =0, end = 0, option = "A") + 
  theme(legend.position = "none", 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") 
g0_freq
ggsave(filename = file.path(figDir, "WeightedSham_Freq_pointWhiskers_BW.pdf"), width =2.5, height = 3)



g0 <- ggplot(pframe[pframe$visitNum_centered %in% c(-25, 25), ], aes(x=treatment, y=predMean))+
    geom_point(aes(color = treatment), alpha = 1) + 
     labs(y = "Sonication frequency (Hz)", x = "Flower treatment") + 
     geom_errorbar(aes(x = treatment, ymin = blo, ymax = bhi, color= treatment), alpha = 1, width = 0.1) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  theme(legend.position = "none", 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") 
g0
ggsave(filename = file.path(figDir, "WeightedSham_Freq_pointWhiskers.pdf"), width =5, height = 3.6)


Fig2a <- g0



## add raw data
g1 <- g0 + geom_point(data = newdf3, aes(x = treatment, y = freq, color = treatment), position = position_jitter(width = 0.1, height = 6), alpha = 0.4, pch = 16, size = 0.2)
g1

ggsave(filename = file.path(figDir, "WeightedSham_Freq_pointWhiskers_rawData.pdf"), width =5, height = 3.6)




# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
g0 <- ggplot(pframe, aes(x=visitNum_centered, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication frequency (Hz)", x = "buzzNumber") + 
     geom_ribbon(aes(x = visitNum_centered, ymin = blo, ymax = bhi), alpha = 0.2) + 
  facet_wrap(~trt2)
g0

# plot effect to see if it agrees
plot(Effect(c("treatment"), m1))

```



### Modeling amplitude with gamm

```{r modeling-amp1}
g01 = gamm4(log(amp_acc) ~ s(visitNum_centered, by = trt2) + trt2 + IT + hive + treatment, random =  ~(1|beeID), data = newdf3)
par(mfrow = c(2,3))
aab <- plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g01$gam) # Summary for paper 
summary(g01$mer)
dev.off()


g011 = gamm4(log(amp_acc) ~ s(visitNum_centered, by = trt2)  + IT + hive + treatment, random =  ~(1|beeID), data = newdf3)
par(mfrow = c(2,3))
aab <- plot(g011$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g011$gam) 
summary(g011$mer)
dev.off()
anova(g01$mer, g011$mer) # go with g011


g02 = gamm4(log(amp_acc) ~ s(visitNum_centered, by = trt2)  + IT  + treatment, random =  ~(1|beeID), data = newdf3)
par(mfrow = c(2,3))
aab <- plot(g02$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g02$gam) # Summary for paper 
summary(g02$mer)
dev.off()

anova(g011$mer, g02$mer) # g02 better


g03 = gamm4(log(amp_acc) ~ s(visitNum_centered, by = trt2)  + treatment, random =  ~(1|beeID), data = newdf3)

anova(g02$mer, g03$mer) #keep g03
summary(g03$mer)

g04 <- gamm4(log(amp_acc) ~ s(visitNum_centered)  + treatment, random =  ~(1|beeID), data = newdf3)

anova(g03$mer, g04$mer) #g04 better
summary(g04$gam) # approx p-value for smooth
summary(g04$mer) 




g44 <- gamm4(log(amp_acc) ~ visitNum_centered  + treatment, random =  ~(1|beeID), data = newdf3)
anova(g04$mer, g44$mer) # g44 better


g05 <- gamm4(log(amp_acc) ~ treatment, random =  ~(1|beeID), data = newdf3)

anova(g44$mer, g05$mer) #g44 better

g06 <- gamm4(log(amp_acc) ~ visitNum_centered , random =  ~(1|beeID), data = newdf3)
anova(g44$mer, g06$mer) #g044 better

# final model -- only treatment matters
ampMod <- g04
summary(ampMod$mer)
summary(ampMod$gam)

plot(ampMod$mer)
par(mfrow = c(2,1))
plot(ampMod$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
dev.off()
```



### Modeling amplitude with lmer

```{r modeling-amp3}
# log transform for acceleration helps model fit better
# start with large model (including interaction)
ma0 = lmer(log(amp_acc) ~ visitNum_centered*trt2 +  hive +IT * treatment+ (1|beeID), data = newdf3, REML = FALSE)
summary(ma0)

ma1 = update(ma0, .~. - visitNum_centered:trt2)

BIC(ma0, ma1) # interaction, ma1, is better

# pval for interaction (in case paper needs it)
anova(ma0, ma1)
summary(ma1)

ma2 <- update(ma1, .~. - trt2)
BIC(ma1, ma2) # keep ma2
summary(ma2)


ma3 <- update(ma2, .~. - hive)
anova(ma2, ma3) # keep ma3

summary(ma3)

ma3.1 <- update(ma3, .~. - visitNum_centered)
anova(ma3, ma3.1) # keep 3.1
summary(ma3.1)

ma4 <- update(ma3.1, .~. - visitNum_centered )
anova(ma3.1, ma4) #keep ma3.1
summary(ma3.1)

m3.2 <- update(ma3.1, .~ . + I(scale(visitNum_centered)) + I(scale(visitNum_centered^2)) + (1|beeID))
summary(m3.2)

anova(ma3.1, m3.2) # keep ma3.1 (according to BIC)



finmod <- update(ma3.1, .~., REML = TRUE)
summary(finmod)

# finmod_centered <- update(ma3.1, .~. - IT*treatment + I(scale(IT, center = TRUE, scale = FALSE))*treatment, REML = TRUE)
# summary(finmod_centered)


# get p-value for interaction for paper
pvalMod <- update(ma3.1, .~.  - IT:treatment)
anova(ma3.1, pvalMod)



# diagnostics
plot(finmod)
qqnorm(ranef(finmod)$beeID[[1]])
qqline(ranef(finmod)$beeID[[1]])


sjp.lmer(finmod) # plot random effects to find any outliers
sjp.lmer(finmod, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects
```


### Bootstrap CI's for amplitude for figure for paper

```{r bootstrap_amp2, cache = TRUE}

# set number of bootstrap replicates for models
nbootSims2 = 10000

# using hive 3, since it's the one with the most data
pframe <- expand.grid(trt2 = levels(droplevels(newdf3$trt2)),
                     IT = c(ITmean, seq(min(newdf3$IT), max(newdf3$IT), length.out = 50)),
                     visitNum_centered = -50:50, 
                     hive = factor(3, levels = levels(newdf3$hive)),  
                     beeID = 99999, 
                     treatment = levels(droplevels(newdf3$treatment)))
pframe$amp_acc <- 0


pframe2 <- pframe[(pframe$trt2 == "Sham first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Sham first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered > 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Weighted" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Weighted first" &
                    pframe$treatment == "Sham" & 
                    pframe$visitNum_centered > 0) 
                    , ]
# exponentiate to put on original scale
pp <- exp(predict(finmod, newdata = pframe2, re.form=NA, type = 'response')) # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(finmod, FUN=function(x) predict(x, pframe2, re.form=NA, type = 'response'), nsim = nbootSims2)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe2$blo<- exp(bb2_se[1,]) # exponentiate to put on original scale
pframe2$bhi<- exp(bb2_se[2,])
pframe2$predMean <- pp
pframe2

### Plot amplitude lmer with trial number

pframe2$trt3 <- mapvalues(pframe2$trt2, from = c("Sham first", "Weighted first"), 
                          to = c("Sham -> Weighted", "Weighted -> Sham"))

g0 <- ggplot(pframe2, aes(x=visitNum_centered, y=predMean))+
     geom_line(aes(color = treatment))+ 
     labs(y = expression ("Sonication acceleration "(m~s^{-2})), x = "Sonication number\n (0 is when treatment switched)") + 
     geom_ribbon(aes(x = visitNum_centered, ymin = blo, ymax = bhi, fill= treatment), alpha = 0.2) + 
  facet_wrap(~trt3) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  theme(legend.position = c(0.5, 0.89), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal")
g0

ggsave(filename = file.path(figDir, "WeightedSham_Timeseries_amp.pdf"), width =5, height = 3.6)



g1 <- g0 + geom_point(data = newdf3[newdf3$visitNum_centered < 50,],
                aes(x = visitNum_centered, y = amp_acc), position = position_jitter(width = 0.5),  size = 0.1, pch = 16, alpha = 0.4) + 
  ylim(c(0, 165)) 
g1

ggsave(filename = file.path(figDir, "WeightedSham_Timeseries_withRawData.pdf"), width =5, height = 3.6)




g1 <- g0 + geom_point(data = newdf3[newdf3$visitNum_centered < 50,],
                aes(x = visitNum_centered, y = amp_acc), position = position_jitter(width = 0.5),  size = 0.1, pch = 16, alpha = 0.4) + 
  scale_y_log10(limits = c(2, 500)) + 
  annotation_logticks(sides = "l", color = 'grey', size = 0.2) 
g1

ggsave(filename = file.path(figDir, "WeightedSham_Timeseries_withRawData_logScale.pdf"), width =5, height = 3.6)

# make figure that isn't faceted by order of treatment

g03 <- ggplot(pframe2, aes(x=visitNum_centered, y=predMean))+
     geom_line(aes(color = treatment))+ 
     labs(y = expression ("Sonication acceleration "(m~s^{-2})), x = "Sonication number\n (0 is when treatment switched)") + 
     geom_ribbon(aes(x = visitNum_centered, ymin = blo, ymax = bhi, fill= treatment), alpha = 0.2) + 
  facet_wrap(~treatment) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  theme(legend.position = c(0.75, 0.8), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="vertical")
  
g03

ggsave(filename = file.path(figDir, "WeightedSham_Timeseries_trtFacet.pdf"), width =5, height = 3.6)


```


### Show amplitude plot without visit number

# look at visit number 1 (right at the time of switching)

```{r}
# mean and 95% CI (without visit number)

g0_amp <- ggplot(pframe2[pframe2$visitNum_centered %in% c(1) & pframe2$IT == ITmean, ], aes(x=treatment, y=predMean))+
    geom_point(aes(color = treatment), alpha = 1) + 
     labs(y = expression ("Sonication acceleration "(m~s^{-2})), x = "Flower treatment") + 
     geom_errorbar(aes(x = treatment, ymin = blo, ymax = bhi, color= treatment), alpha = 1, width = 0.1) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, option = "A", begin =0, end = 0) + 
  theme(legend.position = "none", 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") 
g0_amp
ggsave(filename = file.path(figDir, "WeightedSham_Amp_pointWhiskers_BW.pdf"), width =2.5, height = 3)



g0 <- ggplot(pframe2[pframe2$visitNum_centered %in% c(1) & pframe2$IT == ITmean, ], aes(x=treatment, y=predMean))+
    geom_point(aes(color = treatment), alpha = 1) + 
     labs(y = expression ("Sonication acceleration "(m~s^{-2})), x = "Flower treatment") + 
     geom_errorbar(aes(x = treatment, ymin = blo, ymax = bhi, color= treatment), alpha = 1, width = 0.1) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  theme(legend.position = "none", 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") 
g0
ggsave(filename = file.path(figDir, "WeightedSham_Amp_pointWhiskers.pdf"), width =5, height = 3.6)






# fig 2b

g0IT <- ggplot(pframe2[pframe2$visitNum_centered %in% c(1), ], aes(x=IT, y=predMean))+
    geom_line(aes(color = treatment), alpha = 1) + 
     labs(y = expression ("Sonication acceleration "(m~s^{-2})), x = "Intertegular span (mm)") + 
     geom_ribbon(aes(x = IT, ymin = blo, ymax = bhi, fill = treatment), alpha = 0.2) + 
  scale_color_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Flower treatment", discrete = TRUE, begin =0.3, end = 0.8) + 
  theme(legend.position = "top", 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") 
g0IT
ggsave(g0IT, filename = file.path(figDir, "WeightedSham_Amp_IT_SUPP.pdf"), width =5, height = 3.6)

# make plot with rug
weight_rug <- g0IT + geom_rug(data = newdf3[!duplicated(newdf3$IT),], inherit.aes = FALSE, aes(x = IT))
  
weight_rug
ggsave(weight_rug, filename = file.path(figDir, "WeightedSham_Amp_IT_SUPP_rug.pdf"), width =5, height = 3.6)

Fig2b <- weight_rug



ITmean

g22 <- g0 + 
  annotate(geom="text", x=c(1,2), y=c(0, 0) + 40, label=c("a", "b"),
                color="black")
ggsave(filename = file.path(figDir, "WeightedSham_Amp_pointWhiskers_annot.pdf"), width =5, height = 3.6)

## add raw data
g1 <- g0 + geom_point(data = newdf3, aes(x = treatment, y = amp_acc, color = treatment), position = position_jitter(width = 0.1, height = 6), alpha = 0.4, pch = 16, size = 0.2)
g1

ggsave(filename = file.path(figDir, "WeightedSham_amp_pointWhiskers_rawData.pdf"), width =5, height = 3.6)

```


### Investigate positional amplitude and also compare amplitude from fft 

```{r}
am1 = lmer(log(acc_amplitude_fft_m.s.s) ~ IT + treatment + (1 | beeID) + IT:treatment, data = new_df)
summary(am1)

am2 = update(am1, .~. - IT:treatment)

anova(am1, am2) # BIC says go with smaller model
summary(am2)



amp1 = lmer(log(pos_amplitude_fft_m) ~ IT +treatment + (1 | beeID) , data = new_df)
summary(amp1)
plot(amp1)

ggplot(new_df, aes(x = treatment, y = log(pos_amplitude_fft_m)))+ 
  geom_boxplot()



ggplot(new_df, aes(x = treatment, y = log(amp_acc)))+ 
  geom_boxplot()


ggplot(new_df, aes(x = treatment, y = log(pos_amplitude_fft_m)))+ 
  geom_boxplot()

```














# major break -- below this line is the first method I used


# this is almost identical to the code above



### Modeling frequency with lmer

Use BIC to select model (decide what interactions and covariates to use)

Make sure that REML = FALSE when comparing BIC values

Start with the biggest model of interest, and then see what predictors can be removed

```{r modeling3}

# interaction that may be important, based on domain knowledge
m0 = lmer(freq ~ treatment*IT+  hive + (1|beeID), data = new_df, REML = FALSE)

# no interaction
m1 = lmer(freq ~ treatment+IT+  hive + (1|beeID), data = new_df, REML = FALSE)
BIC(m0, m1) # stay with m1 -- no interaction
anova(m0, m1) # this p-value is very borderline -- we're being 
# conservative to avoid overfitting by using BIC

m2 = lmer(freq ~ hive + IT+ (1|beeID), data = new_df, REML = FALSE)
BIC(m1, m2) # treatment doesn't make the model better

anova(m1, m2) # p-value for paper -- probably don't need to report

# p-value for hive
m3 = lmer(freq ~ treatment + (1|beeID), data = new_df, REML = FALSE)
anova(m1, m3)


# diagnostics -- use REML = TRUE
m1 <- update(m1, .~., REML =TRUE)
summary(m1) # summary for paper
plot(m1)
qqnorm(ranef(m1)$beeID[[1]])
qqline(ranef(m1)$beeID[[1]])

sjp.lmer(m1, type = "re", sort = TRUE) # plot random effects to find any outliers
sjp.lmer(m1, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects
```


### Bootstrap CI's for figure for paper

```{r bootstrap3, cache = TRUE}
# set number of bootstrap replicates for models
nbootSims = 1000

table(new_df$hive) # more trials from hive 3

# using hive 3, since it's the one with the most data
# however, hive doesn't affect model anyway

# calculate an average IT for prediction
ITmean = mean(tapply(new_df$IT, INDEX = new_df$beeID, FUN = function(x) x[1] ))

pframe <- data.frame(treatment = levels(droplevels(new_df$treatment)),
                     IT = ITmean, 
                     hive = factor(3, levels = levels(new_df$hive)),  
                     beeID = 99999)
pframe$freq <- 0
pp <- predict(m1, newdata = pframe, re.form=NA, type = 'response') # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(m1, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<-bb2_se[1,]
pframe$bhi<-bb2_se[2,]
pframe$predMean <- pp
pframe <- pframe[, c('treatment', "blo", "bhi", "predMean")]
pframe
```

### Make frequency plots for paper

```{r figure-making}
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
g0 <- ggplot(pframe, aes(x=treatment, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication frequency (Hz)", x = "Flower treatment") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)
g0
ggsave(plot = g0, filename = file.path(figDir, "HeavyLight_PredsAndCI_freq.pdf"), width =4, height = 3)
```


### Modeling amplitude with lmer

```{r modeling-amp2}
# log transform for acceleration helps model fit better
# start with large model (including interaction)
ma0 = lmer(log(amp_acc) ~ treatment * IT +  hive + (1|beeID), data = new_df, REML = FALSE)

ma1 = lmer(log(amp_acc) ~ treatment + IT +  hive + (1|beeID), data = new_df, REML = FALSE)

BIC(ma0, ma1) # interaction, ma0, is better

# pval for interaction (in case paper needs it)
anova(ma0, ma1)

ma2 <- update(ma0, .~. - hive)
BIC(ma0, ma2) # hive not important

#pval for hive
anova(ma0, ma2)

# fit with reml = TRUE for summarizing
ma3 <- update(ma2,.~., REML = TRUE)
summary(ma3) # final model to report

# diagnostics
plot(ma3)
qqnorm(ranef(ma3)$beeID[[1]])
qqline(ranef(ma3)$beeID[[1]])


sjp.lmer(ma3) # plot random effects to find any outliers
sjp.lmer(ma3, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects
```


### Bootstrap CI's for amplitude for figure for paper

```{r bootstrap_amp3, cache = TRUE}

# set number of bootstrap replicates for models
nbootSims2 = 1000

# using hive 3, since it's the one with the most data
pframe <- data.frame(treatment = levels(droplevels(new_df$treatment)), IT = ITmean, hive = factor(3, levels = levels(new_df$hive)),  beeID = 99999)
pframe$amp_acc <- 0
# exponentiate to put on original scale
pp <- exp(predict(ma3, newdata = pframe, re.form=NA, type = 'response')) # re.form sets all random effects to 0


### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(ma3, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims2)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<- exp(bb2_se[1,]) # exponentiate to put on original scale
pframe$bhi<- exp(bb2_se[2,])
pframe$predMean <- pp
pframe <- pframe[, c('treatment', "blo", "bhi", "predMean")]
pframe
```

### Make amplitude plots for paper

```{r figure-making-amp}
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
# Holding IT = meanIT
print(ITmean)
ga0 <- ggplot(pframe, aes(x=treatment, y=predMean))+
     geom_point()+ 
     labs(y = expression ("Sonication acceleration "(m~s^{-2})), x = "Flower treatment") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1) +
  annotate(geom="text", x=c(1,2), y=c(0, 0) + 40, label=c("a", "b"),
                color="black") 
ga0
ggsave(plot = ga0, filename = file.path(figDir, "HeavyLight_PredsAndCI_amp.pdf"), width =4, height = 3)
```




### Put two plots into a single figure

```{r}

# refref: note I might want to align these two figure for paper (in one plot)
# WeightedSham_Amp_pointWhiskers_BW.pdf
# WeightedSham_Freq_pointWhiskers_BW.pdf

pfreq <- Fig2a

pamp <- Fig2b + theme(legend.position = c(.5,.85)) + ylim(c(10/2,55/2))

  

aa <- plot_grid(pfreq, pamp, align = 'h', axis = 'l', rel_widths = c(1,2))

ggdraw(aa)

ac <- plot_grid(pfreq, pamp, align = 'h', axis = 'l', rel_widths = c(1,2), labels = "AUTO")
ggdraw(ac)

save_plot(file.path(figDir, "CombinedFreqAmp_weightedSham4_noxlab_IT.svg"), plot = aa, base_width = 7, base_height = 3)

save_plot(file.path(figDir, "CombinedFreqAmp_weightedSham4_annot_IT.svg"), plot = ac, base_width = 7, base_height = 3)

```


```{r}
sessionInfo()
````

