---
output: html_document
editor_options: 
  chunk_output_type: console
---

<pre>
## Visualize experiment from hives 7-8
## Bees were given two treatments -- reward first or second
</pre>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/cswitzer/Dropbox/SonicationBehavior")
```

```{r}
#install packages
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("ggplot2", "reshape2", 'lme4', 'sjPlot', "multcomp", 
              "plyr", "effects", "viridis", "GGally", "survminer", 
              "tidyr","gamm4")
ipak(packages)

# set ggplot theme
theme_set(theme_classic() + theme(axis.text=element_text(colour="black")))

# define data and figure directories
dataDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehData"
figDir <- "/Users/cswitzer/Dropbox/SonicationBehavior/SonBehFigs"


# print system info
print(paste("last run ", Sys.time()))
print(R.version)
```


# Read in data
```{r}
bees <- read.csv(file.path(dataDir, "03_StartStop_cleaned.csv"))
bees$hive = as.factor(bees$hive)
head(bees)

# calculate mean freq rewarded vs. unrewarded for each bee
# to make sure it agrees with what is already in the dataframe
tapply(bees$freq, INDEX = list(bees$beeID, bees$treatment_rewarded), mean)

# number of bees -- should be 96
length(unique(bees$BeeNumCol))

# num observations -- should be 5851
nrow(bees)
```

## Visualize data
```{r}
ggplot(bees, aes(y = freq, x = treatment_rewarded)) +
     geom_boxplot() +  
     stat_smooth(method = "loess", span = 0.75)

bees$trt2 <- mapvalues(bees$treatment, from = c("RewFir", "RewSec"), 
                       to = c("Rewarded -> Unrewarded", "Unrewarded -> Rewarded"))

bees$visitNum_centered <- bees$visitNum - 50

ggplot(bees, aes(x = visitNum_centered, y = freq, color = treatment_rewarded)) +
     stat_smooth(method = "loess", span = 1) + 
     xlab("Sonication number\n (0 is when treatment switched)") + 
     ylab("Sonication Frequency(hz)") + 
     facet_wrap(~trt2) + 
     scale_color_viridis(name = "Rewarded Buzzes", 
                         discrete = TRUE, begin = 0.2, end = 0.6) 
#ggsave("beeRewards.png", width = 8, height = 5)


# make survival plot
survBees <- t(sapply(unique(bees$beeID), function(x) {
     tmp <- bees[bees$beeID == x, ]
     return(tmp[which.max(tmp$visitNum), ])
}))


lastObs <- as.data.frame(survBees)
lastObs$index <- as.numeric(lastObs$visitNum)


# 1 = censored, 2 = stopped buzzing for 5 min
lastObs$status <- mapvalues(lastObs$index == 100, from = c(TRUE, FALSE), to = c(1, 2))

# create survival object
lastObs$survObj <- with(lastObs, Surv(index, status == 2))
head(lastObs)

lastObs$treatment <- as.factor(unlist(lastObs$treatment))


## Kaplan-Meier estimator. 
km.as.one <- survfit(survObj ~ 1, data = lastObs, conf.type = "log-log")
km.by.trt <- survfit(survObj ~ treatment, data = lastObs, conf.type = "log-log")

par(mfrow = c(1,1))
plot(km.by.trt)

gsv <- ggsurvplot(km.by.trt, CI = FALSE) 

gsv$plot + facet_wrap(~treatment) + 
     theme_bw()   + 
     theme(legend.position = "none")  + 
     scale_color_viridis(discrete = TRUE, begin = 0.3, end = 0.7, option = "magma")  + 
     xlab("Buzz Number") + 
     ylab("Proportion Sonicating") 

#ggsave("beeSonicationPercentage.pdf", width = 8, height = 5)


ggsurv(km.by.trt, CI = FALSE, surv.col = c("#a6cee3", "#1f78b4")) + 
     ylim(c(0, 1)) + 
     xlab("Buzz Number") + 
     ylab("Proportion Sonicating") 

xtabs(~lastObs$treatment + as.character(lastObs$hive))

```


### Visualize data -- refref: averaged by bee and trial

This is not super informative

```{r}
beeAvgs = bees[bees$visitNum == 1, ]
nrow(beeAvgs) # should be 96
head(beeAvgs)


data_long <- gather(beeAvgs[, c("BeeNumCol", "meanFreqRewarded", "meanFreqUnrewarded")], condition, meanFreq, meanFreqRewarded, meanFreqUnrewarded)
head(data_long)


ggplot(data_long, aes(x = condition, y = meanFreq)) + 
  geom_boxplot()

# look at diffs
beeAvgs$buzzDiff = beeAvgs$meanFreqRewarded - beeAvgs$meanFreqUnrewarded
hist(beeAvgs$buzzDiff)
```




# Analysis with start/stop trials

#### GAMM4 shows time series

```{r}

# start with gamm so I can show change by visit number
g00 = gamm4(log(amp_acc) ~ s(visitNum_centered, by = trt2) + treatment_rewarded * IT_mm + hive , random =  ~(1|beeID), data = bees)
par(mfrow = c(2,3))
aab <- plot(g00$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g00$gam) # Summary for paper 
summary(g00$mer)
dev.off()

summary(g00$mer)
summary(g00$gam)

g01 = gamm4(log(amp_acc) ~ s(visitNum_centered) + treatment_rewarded * IT_mm + hive , random =  ~(1|beeID), data = bees)
anova(g00$mer, g01$mer) # stay with g01
summary(g01$mer)
summary(g01$gam)

g03 <- gamm4(log(amp_acc) ~  treatment_rewarded * IT_mm + hive , random =  ~(1|beeID), data = bees)
anova(g01$mer, g03$mer) # g03 better
summary(g03$mer) # stay with g03


g04 <- gamm4(log(amp_acc) ~  treatment_rewarded + I(scale(IT_mm)) + hive , random =  ~(1|beeID), data = bees)
anova(g03$mer, g04$mer) # g03 better
summary(g03$mer) # stay with g03

# refit with lmer



g03 <- lmer(log(amp_acc) ~  treatment_rewarded + IT_mm + hive + (1|beeID), data = bees, REML = TRUE)
summary(g03) # summary for paper

finmod <- g03


## predict gamm

# using hive 7, since it's the one with the most data

# calculate an average IT for prediction
ITmean = mean(tapply(bees$IT_mm, INDEX = bees$beeID, FUN = function(x) x[1] ))
pframe = expand.grid(trt2 = levels(droplevels(bees$trt2)), 
                     visitNum_centered = -50:50, 
                     IT_mm = ITmean,
                     hive = factor(7, levels = levels(bees$hive)), 
                     beeID = 99999, 
                     treatment_rewarded = levels(bees$treatment_rewarded))

pframe$amp_acc <- 0

pframe <- pframe[(pframe$trt2 == "Rewarded -> Unrewarded" &
                    pframe$treatment_rewarded == " F" & 
                    pframe$visitNum_centered > 0) |
                    (pframe$trt2 == "Rewarded -> Unrewarded" &
                    pframe$treatment_rewarded == " T" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Unrewarded -> Rewarded" &
                    pframe$treatment_rewarded == " F" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Unrewarded -> Rewarded" &
                    pframe$treatment_rewarded == " T" & 
                    pframe$visitNum_centered > 0) 
                    , ]

# predict make plots, using full model (g00)

pp <- exp(predict(finmod, newdata = pframe, re.form=NA, type = 'response')) # re.form sets all random effects to 0

nbootSims2 = 1000
### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(finmod, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims2)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<- exp(bb2_se[1,]) # exponentiate to put on original scale
pframe$bhi<- exp(bb2_se[2,])
pframe$predMean <- pp
pframe

# Holding IT = meanIT
print(ITmean)
ga0 <- ggplot(pframe, aes(x=treatment_rewarded, y=predMean))+
     geom_point()+ 
     labs(y = expression ("Sonication amplitude "(m~s^{-2})), x = "Rewarded") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1) +
  annotate(geom="text", x=c(1,2), y=c(0, 0) + 60, label=c("a", "b"),
                color="black") 
ga0
ggsave(plot = ga0, filename = file.path(figDir, "StartStopBootAmp.pdf"), width =4, height = 3)



pframe$treatment_rewarded2 <- mapvalues(pframe$treatment_rewarded, from = c(" F", " T"), 
                                        to = c("No", "Yes"))

pframe$treatment_rewarded2 <- relevel(pframe$treatment_rewarded2, ref = "Yes")

g03 <- ggplot(pframe, aes(x=visitNum_centered, y = predMean))   + 
  geom_line(aes(color = treatment_rewarded2)) + 
  # geom_line(aes(y=exp(fit-2*se), col=factor(group)), linetype="dashed") + 
  # geom_line(aes(y=exp(fit+2*se), col=factor(group)), linetype="dashed") + 
  geom_ribbon(aes(x=as.numeric(visitNum_centered), ymax=bhi, ymin=blo, fill = treatment_rewarded2), 
              alpha=0.2) +
  theme(legend.position = c(0.5, 0.85), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") + 
  facet_wrap(~trt2) + 
  labs(y = expression ("Sonication amplitude "(m~s^{-2})), 
       x = "Sonication number\n (0 is when treatment switched)") +
  scale_color_viridis(name = "Bee rewarded", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Bee rewarded", discrete = TRUE, begin =0.3, end = 0.8)  + 
  ylim(c(42, 57))
  # geom_segment(aes(y = 40, yend = 55, x = 0, xend = 0), inherit.aes = FALSE, lty = 2, alpha = 0.2, lwd = 0.2)
  # geom_vline(aes(xintercept = 50, ymin = 40, ymax = 55), lty = 2, alpha = 0.2)
g03
ggsave(filename = file.path(figDir, "LMER_amplitude_visitNum.pdf"), width =5, height = 3.6)

```



```{r, eval = FALSE}


# old code, from when I though gamm was the way

plot.df = cbind(pframe, as.data.frame(predict( g03, pframe, se.fit = T)))
pred.gamm.plot = plot.df
pred.gamm.plot$group = pred.gamm.plot$trt2
pred.gamm.plot$se = pred.gamm.plot$se.fit
bees$prediction <- as.vector(predict.gam(g04$gam))

pred.gamm.plot$rewarded = mapvalues(pred.gamm.plot$treatment_rewarded, from = c(" T", " F"), to = c("Yes", "No"))
pred.gamm.plot$rewarded <- relevel(pred.gamm.plot$rewarded, ref = "Yes")

g0 <- ggplot(pred.gamm.plot, aes(x=visitNum_centered, y = exp(as.numeric(fit))))   + 
  geom_line(aes(color = rewarded)) + 
  # geom_line(aes(y=exp(fit-2*se), col=factor(group)), linetype="dashed") + 
  # geom_line(aes(y=exp(fit+2*se), col=factor(group)), linetype="dashed") + 
  geom_ribbon(aes(x=as.numeric(visitNum_centered), ymax=exp(fit+2*se), ymin=exp(fit-2*se), fill = rewarded), 
              alpha=0.2) +
  theme(legend.position = c(0.754, 0.85), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") + 
  facet_wrap(~group) + 
  labs(y = expression ("Sonication amplitude "(m~s^{-2})), 
       x = "Sonication number\n (0 is when treatment switched)") +
  scale_color_viridis(name = "Bee rewarded", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Bee rewarded", discrete = TRUE, begin =0.3, end = 0.8) 
  # geom_segment(aes(y = 40, yend = 55, x = 0, xend = 0), inherit.aes = FALSE, lty = 2, alpha = 0.2, lwd = 0.2)
  # geom_vline(aes(xintercept = 50, ymin = 40, ymax = 55), lty = 2, alpha = 0.2)
g0
# ggsave(filename = file.path(figDir, "StartStopPreds_amp_gamm2.pdf"), width =5, height = 3.6)


# add raw data to plot
g1 <- g0 + geom_point(data = bees, aes(x = visitNum_centered, y = amp_acc), 
                      size = 0.1, pch = 16, 
                      alpha = 0.1, position = position_jitter(width = 0.6)) + 
  ylim(c(0, 200)) + 
  theme(legend.position = c(0.5, 0.89), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal")
g1

# ggsave(filename = file.path(figDir, "StartStopPreds_amp_gamm2_rawData.pdf"), width =5, height = 3.6)




# add raw data to plot
g2 <- g0 + geom_point(data = bees, aes(x = visitNum_centered, y = amp_acc), 
                      size = 0.1, pch = 16, 
                      alpha = 0.1, position = position_jitter(width = 0.6)) + 
  ylim(c(0, 200)) + 
  theme(legend.position = c(0.5, 0.89), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") + 
  scale_y_log10(limits = c(2, 500)) + 
  annotation_logticks(sides = "l", color = 'grey', size = 0.2) 
g2

# ggsave(filename = file.path(figDir, "StartStopPreds_amp_gamm2_rawData_logScale.pdf"), width =5, height = 3.6)



# plot with different color scheme
ggplot(pred.gamm.plot, aes(x=as.numeric(visitNum_centered), y = exp(as.numeric(fit))))   + 
  geom_line(aes(color = group)) + 
  # geom_line(aes(y=exp(fit-2*se), col=factor(group)), linetype="dashed") + 
  # geom_line(aes(y=exp(fit+2*se), col=factor(group)), linetype="dashed") + 
  geom_ribbon(aes(x=as.numeric(visitNum_centered), ymax=exp(fit+2*se), ymin=exp(fit-2*se), fill = group), 
              alpha=0.2) +
  theme(legend.position = "none") + 
  facet_wrap(~group) + 
  labs(y = expression ("Sonication amplitude "(m~s^{-2})), x = "Visit Number") +
  scale_color_viridis(discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(discrete = TRUE, begin =0.3, end = 0.8)+ 
  geom_vline(aes(xintercept = 50), lty = 2, alpha = 0.2)


# plot raw data with estimated smooths (from gam)

# plot raw data with loess
ggplot(bees[bees$totalBuzzesNotRewarded > 30 & bees$totalRewards > 30, ], 
       aes(x=as.numeric(visitNum), y = amp_acc))   + 
  geom_point(alpha = 0.1, size = 0.1) + 
   # geom_line(aes(colour = as.factor(beeID)), alpha = 0.2) +
  # geom_line(stat = "smooth", method = "lm", aes(colour = as.factor(beeID)), alpha = 0.2) +
  geom_smooth(aes(group = trt2), method = 'auto') +
  theme(legend.position = "none") + 
  facet_wrap(~trt2) + 
  labs(y = expression ("Sonication amplitude "(m~s^{-2})), x = "Visit Number") +
  scale_color_viridis(discrete = TRUE) + 
  scale_fill_viridis(discrete = TRUE, end = 0.8)+ 
  geom_vline(aes(xintercept = 50), lty = 2, alpha = 0.2)

```




### GAMM for freq

```{r}

# start with gamm so I can show change by visit number
g01 = gamm4(freq ~ s(visitNum, by = trt2) +  trt2 + IT_mm*treatment_rewarded + hive , random =  ~(1|beeID), data = bees)
par(mfrow = c(2,3))
aab <- plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE)
summary(g01$gam) # Summary for paper 
summary(g01$mer)
dev.off()

g02 = gamm4(freq ~ s(visitNum) +  trt2 + IT_mm*treatment_rewarded + hive , random =  ~(1|beeID), data = bees)
anova(g01$mer, g02$mer) # go with g02
summary(g02$mer)

g03 <- gamm4(freq ~  trt2 + IT_mm*treatment_rewarded + hive , random =  ~(1|beeID), data = bees)
anova(g02$mer, g03$mer) # go with g03
summary(g03$mer)

g04 <- gamm4(freq ~ IT_mm*treatment_rewarded + hive , random =  ~(1|beeID), data = bees)
anova(g03$mer, g04$mer) # g04
summary(g04$mer)

finmod <- lmer(freq ~ IT_mm*treatment_rewarded + hive + (1|beeID), data = bees)
summary(finmod)

## refref here: whittle down GAMM like amplitude above

finmod <- g03
plot(finmod)

## refref diagnostics for this model


## predict gamm

# using hive 7, since it's the one with the most data

# calculate an average IT for prediction
ITmean = mean(tapply(bees$IT_mm, INDEX = bees$beeID, FUN = function(x) x[1] ))
pframe = expand.grid(trt2 = levels(droplevels(bees$trt2)), 
                     visitNum_centered = -50:50, 
                     IT_mm = ITmean,
                     hive = factor(7, levels = levels(bees$hive)), 
                     beeID = 99999, 
                     treatment_rewarded = levels(bees$treatment_rewarded))

pframe$freq <- 0

pframe <- pframe[(pframe$trt2 == "Rewarded -> Unrewarded" &
                    pframe$treatment_rewarded == " F" & 
                    pframe$visitNum_centered > 0) |
                    (pframe$trt2 == "Rewarded -> Unrewarded" &
                    pframe$treatment_rewarded == " T" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Unrewarded -> Rewarded" &
                    pframe$treatment_rewarded == " F" & 
                    pframe$visitNum_centered < 0) |
                    (pframe$trt2 == "Unrewarded -> Rewarded" &
                    pframe$treatment_rewarded == " T" & 
                    pframe$visitNum_centered > 0) 
                    , ]

# predict make plots, using full model 

pp <- predict(finmod, newdata = pframe, re.form=NA, type = 'response') # re.form sets all random effects to 0

nbootSims2 = 10
### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(finmod, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims2)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<- (bb2_se[1,]) # exponentiate to put on original scale
pframe$bhi<- (bb2_se[2,])
pframe$predMean <- pp
pframe

# Holding IT = meanIT
print(ITmean)
ga0 <- ggplot(pframe, aes(x=treatment_rewarded, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication Frequency (Hz)", x = "Rewarded") + 
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1)
ga0
ggsave(plot = ga0, filename = file.path(figDir, "StartStop_Freq.pdf"), width =4, height = 3)



pframe$treatment_rewarded2 <- mapvalues(pframe$treatment_rewarded, from = c(" F", " T"), 
                                        to = c("No", "Yes"))

pframe$treatment_rewarded2 <- relevel(pframe$treatment_rewarded2, ref = "Yes")

g03 <- ggplot(pframe, aes(x=visitNum_centered, y = predMean))   + 
  geom_line(aes(color = treatment_rewarded2)) + 
  # geom_line(aes(y=exp(fit-2*se), col=factor(group)), linetype="dashed") + 
  # geom_line(aes(y=exp(fit+2*se), col=factor(group)), linetype="dashed") + 
  geom_ribbon(aes(x=as.numeric(visitNum_centered), ymax=bhi, ymin=blo, fill = treatment_rewarded2), 
              alpha=0.2) +
  theme(legend.position = c(0.5, 0.85), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") + 
  facet_wrap(~trt2) + 
  labs(y = "Sonication Frequency (Hz)", 
       x = "Sonication number\n (0 is when treatment switched)") +
  scale_color_viridis(name = "Bee rewarded", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Bee rewarded", discrete = TRUE, begin =0.3, end = 0.8)
  # geom_segment(aes(y = 40, yend = 55, x = 0, xend = 0), inherit.aes = FALSE, lty = 2, alpha = 0.2, lwd = 0.2)
  # geom_vline(aes(xintercept = 50, ymin = 40, ymax = 55), lty = 2, alpha = 0.2)
g03
# ggsave(filename = file.path(figDir, "LMER_amplitude_visitNum.pdf"), width =5, height = 3.6)

```



```



```{r, eval = FALSE}


# old analysis that was for GAMM
pp <- predict(g01$gam, newdata = pframe2, re.form=NA, type = 'response') # re.form sets all random effects to 0


plot.df = cbind(pframe2, as.data.frame(predict( g01$gam, pframe2, se.fit = T)))
pred.gamm.plot = plot.df
pred.gamm.plot$group = pred.gamm.plot$trt2
pred.gamm.plot$se = pred.gamm.plot$se.fit
bees$prediction <- as.vector(predict.gam(g01$gam))

pred.gamm.plot$rewarded = mapvalues(pred.gamm.plot$treatment_rewarded, from = c(" T", " F"), to = c("Yes", "No"))

pred.gamm.plot$rewarded <- relevel(pred.gamm.plot$rewarded, ref = "Yes")


ggplot(pred.gamm.plot, aes(x=as.numeric(visitNum), y = (as.numeric(fit))))   + 
  geom_line(aes(color = rewarded)) + 
  # geom_line(aes(y=exp(fit-2*se), col=factor(group)), linetype="dashed") + 
  # geom_line(aes(y=exp(fit+2*se), col=factor(group)), linetype="dashed") + 
  geom_ribbon(aes(x=as.numeric(visitNum), ymax=(fit+2*se), ymin=(fit-2*se), fill = rewarded), 
              alpha=0.2) +
  theme(legend.position = c(0.754, 0.85), 
        legend.background = element_rect(fill=alpha('gray95', 1)), 
        legend.direction="horizontal") + 
  facet_wrap(~group) + 
  labs(y = "Sonication frequency (Hz)", x = "Visit Number") + 
  scale_color_viridis(name = "Bee rewarded", discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(name = "Bee rewarded", discrete = TRUE, begin =0.3, end = 0.8)+ 
  geom_segment(aes(y = 315, yend = 340, x = 50, xend = 50), inherit.aes = FALSE, lty = 2, alpha = 0.2, lwd = 0.2)
  # geom_vline(aes(xintercept = 50, ymin = 40, ymax = 55), lty = 2, alpha = 0.2)
ggsave(filename = file.path(figDir, "StartStopPreds_freq_gamm.pdf"), width =5, height = 3.6)




ggplot(pred.gamm.plot, aes(x=as.numeric(visitNum), y = (as.numeric(fit))))   + 
  geom_line(aes(color = group)) + 
  geom_ribbon(aes(x=as.numeric(visitNum), ymax=(fit+2*se), ymin=(fit-2*se), fill = group), 
              alpha=0.2) +
  theme(legend.position = "none") + 
  facet_wrap(~group) + 
  labs(y = "Sonication frequency (Hz)", x = "Visit Number") + 
  scale_color_viridis(discrete = TRUE, begin =0.3, end = 0.8) + 
  scale_fill_viridis(discrete = TRUE, begin =0.3, end = 0.8) + 
  geom_vline(aes(xintercept = 50), lty = 2, alpha = 0.2)

```

##### LMER for amp_acc

```{r}
head(bees)

mod0 <- lmer(log(amp_acc) ~ trt2*visitNum + trt2*I(scale(visitNum)^2) + trt2*I(scale(visitNum)^3) + IT_mm + hive + (1|beeID), data = bees, REML = FALSE)
summary(mod0)


mod1 <- update(mod0, .~. - trt2:I(scale(visitNum)^3))

anova(mod0, mod1) # BIC says to go with mod1
summary(mod1)



mod2 <- update(mod1, .~. -  trt2:I(scale(visitNum)^2))
anova(mod2, mod1) # bic says go with mod 2

summary(mod2)

mod3 <- update(mod2, .~. - I(scale(visitNum)^3))
anova(mod3, mod2)
summary(mod3)

mod4 <- update(mod3, .~. - I(scale(visitNum)^2))
anova(mod3, mod4)

summary(mod4)
mod5 <- update(mod4, .~. - trt2:visitNum)

anova(mod4, mod5)  # go with mod4
summary(mod5)

mod0 <- update(mod4, .~., REML = TRUE)
summary(mod0)


# predict mod0 
# calculate an average IT for prediction
ITmean = mean(tapply(bees$IT_mm, INDEX = bees$beeID, FUN = function(x) x[1] ))
pframe = expand.grid(trt2 = levels(droplevels(bees$trt2)), 
                     visitNum = 1:100, 
                     treatment_rewarded = levels(bees$treatment_rewarded),
                     IT_mm = ITmean,
                     hive = factor(7, levels = levels(bees$hive)), beeID = 99999)

pframe$freq <- 0

pframe2 = pframe[(pframe$trt2 == "Rewarded 1st 50 Buzzes" & 
                    pframe$treatment_rewarded == " T" & 
                   pframe$visitNum < 50) |
                   (pframe$trt2 == "Rewarded 2nd 50 Buzzes" & 
                    pframe$treatment_rewarded == " T" & 
                   pframe$visitNum > 50) |
                   (pframe$trt2 == "Rewarded 1st 50 Buzzes" & 
                    pframe$treatment_rewarded == " F" & 
                   pframe$visitNum > 50) |
                   (pframe$trt2 == "Rewarded 2nd 50 Buzzes" & 
                    pframe$treatment_rewarded == " F" & 
                   pframe$visitNum < 50), ]

pp <- predict(mod0, newdata = pframe2, re.form=NA, type = 'response') # re.form sets all random effects to 0



### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(mod0, FUN=function(x) predict(x, pframe2, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe2$blo<-bb2_se[1,]
pframe2$bhi<-bb2_se[2,]
pframe2$predMean <- pp
pframe2


ggplot(pframe2, aes(x=visitNum, color = treatment_rewarded, y=predMean))+
     geom_line()+ 
     labs(y = "Sonication amplitude", x = "Sonication Number") + 
    theme(legend.position = c(.74,.8), 
          legend.background = element_rect(fill=alpha('gray90', 0.5))) +
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1, lwd = 1, alpha = 0.2) + 
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded", begin =0.3, end = 0.8) +
  facet_wrap(~trt2) 


```


### LMER analysis for frequency

#### refref: todo -- more model selection and diagnostics 
 -- get p-vals, maybe?

```{r}
mod1 <- lmer(freq ~ trt2 * I(scale(visitNum)) + I(scale(visitNum)^2) + trt2 * I(scale(visitNum)^2) + trt2 * I(scale(visitNum)^3) + I(scale(visitNum)^3) + treatment_rewarded + IT_mm + hive + (1|beeID), REML = FALSE, data = bees)

mod1_1 <- update(mod1, .~. - trt2 : I(scale(visitNum)^3))
BIC(mod1, mod1_1) # mod 1_1 better

summary(mod1_1)


mod1_2 <- update(mod1_1, .~. - trt2 : I(scale(visitNum)^2))
BIC(mod1_2, mod1_1) # mod 1_2 better

summary(mod1_2)


mod1_3 <- update(mod1_2, .~.- trt2 : I(scale(visitNum)))
BIC(mod1_2, mod1_3) # mod 1_3 better

summary(mod1_3)

mod1_4 <- update(mod1_3, .~.- trt2 )
BIC(mod1_4, mod1_3) # mod 1_4 better

summary(mod1_4)

mod1_5 <- update(mod1_4, .~. - IT_mm)
BIC(mod1_4, mod1_5) # mod 1_5 better
summary(mod1_5)



mod2 <- update(mod1_5, .~., REML = TRUE)

vif.mer <- function (fit) {
    ## adapted from rms::vif

    v <- vcov(fit)
    nam <- names(fixef(fit))

    ## exclude intercepts
    ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
    if (ns > 0) {
        v <- v[-(1:ns), -(1:ns), drop = FALSE]
        nam <- nam[-(1:ns)]
    }

    d <- diag(v)^0.5
    v <- diag(solve(v/(d %o% d)))
    names(v) <- nam
    v
}

vif.mer(mod2)

summary(mod2)

BIC(mod1, mod2) # drop interaction
anova(mod1, mod2) # drop interaction


# diagnostics -- use REML = TRUE
m1 <- update(mod2, .~., REML =TRUE)
summary(m1) # summary for paper

vif.mer(m1)




plot(m1)
qqnorm(ranef(m1)$beeID[[1]])
qqline(ranef(m1)$beeID[[1]])

sjp.lmer(m1, type = "re", sort = TRUE) # plot random effects to find any outliers
sjp.lmer(m1, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects
```


### Bootstrap CI's for figure for paper

```{r bootstrap, cache = TRUE}
# set number of bootstrap replicates for models
nbootSims = 10

table(bees$hive) # more trials from hive 7

# using hive 7, since it's the one with the most data

# calculate an average IT for prediction
ITmean = mean(tapply(bees$IT_mm, INDEX = bees$beeID, FUN = function(x) x[1] ))
pframe = expand.grid(trt2 = levels(droplevels(bees$trt2)), 
                     visitNum = 1:100, 
                     treatment_rewarded = levels(bees$treatment_rewarded),
                     IT_mm = ITmean,
                     hive = factor(7, levels = levels(bees$hive)), beeID = 99999)

pframe$freq <- 0

pframe2 = pframe[(pframe$trt2 == "Rewarded 1st 50 Buzzes" & 
                    pframe$treatment_rewarded == " T" & 
                   pframe$visitNum < 50) |
                   (pframe$trt2 == "Rewarded 2nd 50 Buzzes" & 
                    pframe$treatment_rewarded == " T" & 
                   pframe$visitNum > 50) |
                   (pframe$trt2 == "Rewarded 1st 50 Buzzes" & 
                    pframe$treatment_rewarded == " F" & 
                   pframe$visitNum > 50) |
                   (pframe$trt2 == "Rewarded 2nd 50 Buzzes" & 
                    pframe$treatment_rewarded == " F" & 
                   pframe$visitNum < 50), ]

pp <- predict(m1, newdata = pframe2, re.form=NA, type = 'response') # re.form sets all random effects to 0



### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(m1, FUN=function(x) predict(x, pframe2, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe2$blo<-bb2_se[1,]
pframe2$bhi<-bb2_se[2,]
pframe2$predMean <- pp
pframe2
```


### Make frequency plots for paper -- with visitNum

```{r figure-making1}
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"

ggplot(pframe2, aes(x=visitNum, color = treatment_rewarded, y=predMean))+
     geom_line()+ 
     labs(y = "Sonication frequency (Hz)", x = "Sonication Number") + 
    theme(legend.position = c(.74,.8), 
          legend.background = element_rect(fill=alpha('gray90', 0.5))) +
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1, lwd = 1, alpha = 0.2) + 
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded", begin =0.3, end = 0.8) +
  facet_wrap(~trt2) 

g0 <- ggplot(pframe, aes(x=trt2, color = treatment_rewarded, y=predMean))+
     geom_point(position = position_dodge(width = 0.4))+ 
     labs(y = "Sonication frequency (Hz)", x = "Treatment") + 
    theme(legend.position = c(.74,.8), 
          legend.background = element_rect(fill=alpha('gray90', 0.5))) +
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1, position = position_dodge(width = 0.4)) + 
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded", begin =0.3, end = 0.8) + 
  geom_vline(aes(xintercept = 1.5))
g0
ggsave(plot = g0, filename = file.path(figDir, "StartStopPreds_freq.pdf"), width =4, height = 3)
```





### Make frequency plots for paper

```{r figure-making}
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
facLevs = c("Rewarded 1st 50 Buzzes\n (buzz #51-100)", 
               "Rewarded 2nd 50 Buzzes\n (buzz #1-50)", 
               "Rewarded 1st 50 Buzzes\n (buzz #1-50)",
               "Rewarded 2nd 50 Buzzes\n (buzz #50-100)"
               )

pframe$t2 <- factor(facLevs,  levels = facLevs[c(3,1,2,4)])


g00 <- ggplot(pframe, aes(x=t2, color = treatment_rewarded, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication frequency (Hz)", x = "Treatment") + 
    theme(legend.position = c(.74,.8), 
          legend.background = element_rect(fill=alpha('gray90', 0.5)), 
          axis.text.x = element_text(angle = 25, hjust = 0.9)) +
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1) + 
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded?", begin =0.3, end = 0.8) + 
  geom_vline(aes(xintercept = 2.5))
g00
ggsave(plot = g00, filename = file.path(figDir, "StartStopPreds_freq2.pdf"), width =5, height = 3.5)


fac2Levs = c("51-100","1-50", "1-50","51-100")
              

pframe$buzzNums <- factor(fac2Levs,  levels = fac2Levs[c(3,1)])
pframe$beesRewarded = mapvalues(pframe$treatment_rewarded, from  = c(" F", " T"), 
                                to = c("No", "Yes"))

g001 <- ggplot(pframe, aes(x=buzzNums, color = beesRewarded, y=predMean))+
     geom_point()+ 
     labs(y = "Sonication frequency (Hz)", x = "Sonication Number") + 
    theme(legend.position = c(.74,.8), 
          legend.background = element_rect(fill=alpha('gray90', 0.5))) +
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1) + 
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded", begin =0.3, end = 0.8) +
  facet_wrap(~trt2) + 
  ylim(c(315, 357))
g001
ggsave(plot = g001, filename = file.path(figDir, "StartStopPreds_freq3.pdf"), width =5, height = 3.5)

g0 <- ggplot(pframe, aes(x=trt2, color = treatment_rewarded, y=predMean))+
     geom_point(position = position_dodge(width = 0.4))+ 
     labs(y = "Sonication frequency (Hz)", x = "Treatment") + 
    theme(legend.position = c(.74,.8), 
          legend.background = element_rect(fill=alpha('gray90', 0.5))) +
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1, position = position_dodge(width = 0.4)) + 
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded", begin =0.3, end = 0.8) + 
  geom_vline(aes(xintercept = 1.5))
g0
ggsave(plot = g0, filename = file.path(figDir, "StartStopPreds_freq.pdf"), width =4, height = 3)
```

### Amplitude analysis for paper

```{r}
head(bees)

mod0 <- lmer(log(amp_acc) ~ trt2*visitNum + IT_mm + hive + (1|beeID), data = bees, REML = FALSE)
mod1 <- lmer(log(amp_acc) ~ trt2+ visitNum + IT_mm + hive + (1|beeID), data = bees, REML = FALSE)

mod0 <- update(mod0, .~., REML = TRUE)
summary(mod0)

plot(mod0)

BIC(mod1, mod0) # keep mod1 (no interaction)
anova(mod0, mod1) #keep mod1

mod2 <- update(mod1, .~. - trt2)
BIC(mod1, mod2) # keep mod2
anova(mod1, mod2) # disagrees with BIC -- barely


m2 <- update(mod2, .~., REML = TRUE)


summary(m2) # summary for paper

# pval for paper
m22 <- update(m2, .~. - treatment_rewarded , REML = TRUE)
anova(m2, m22) # p-val for interaction

plot(m2)
qqnorm(ranef(m2)$beeID[[1]])
qqline(ranef(m2)$beeID[[1]])

sjp.lmer(mod0, type = "re", sort = TRUE) # plot random effects to find any outliers
sjp.lmer(mod0, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects

```




# GLMER

```{r}
mod0 <- glmer(amp_acc ~ trt2 * treatment_rewarded + scale(IT_mm) + hive + (1|beeID), family = Gamma("log"), data = bees)
summary(mod0)

mod0 <- glmer(amp_acc ~ treatment_rewarded + scale(IT_mm) + hive + (1|beeID), family = Gamma("log"), data = bees)
summary(mod0)


BIC(mod1, mod0) # keep mod1
anova(mod0, mod1) #keep mod1

mod2 <- update(mod1, .~. - trt2)
BIC(mod1, mod2) # keep mod2
anova(mod1, mod2) # disagrees with BIC -- barely
summary(mod1)


m2 <- update(mod2, .~., REML = TRUE)


summary(m2) # summary for paper

# pval for paper
m22 <- update(m2, .~. - treatment_rewarded , REML = TRUE)
anova(m2, m22) # p-val for interaction

plot(m2)
qqnorm(ranef(m2)$beeID[[1]])
qqline(ranef(m2)$beeID[[1]])

sjp.lmer(m2, type = "re", sort = TRUE) # plot random effects to find any outliers
sjp.lmer(m2, type = 'fe', sort = TRUE, p.kr = FALSE) # plot fixed effects
```



### Bootstrap Amp CI's for figure for paper

```{r bootstrap_Amp, cache = TRUE}
# set number of bootstrap replicates for models
nbootSims = 100

table(bees$hive) # more trials from hive 3

# using hive 7, since it's the one with the most data

pframe <- data.frame(trt2 = rep(levels(droplevels(bees$trt2)), 2),
                     treatment_rewarded = rep(levels(bees$treatment_rewarded), each = 2),
                     IT_mm = ITmean, 
                     hive = factor(7, levels = levels(bees$hive)),  
                     beeID = 99999)
pframe$amp_acc <- 0

m00 <- update(mod0, .~., REML = TRUE)
pp <- exp(predict(m00, newdata = pframe, re.form=NA, type = 'response')) # re.form sets all random effects to 0



### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(m00, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<-exp(bb2_se[1,])
pframe$bhi<-exp(bb2_se[2,])
pframe$predMean <- pp
pframe
```


#### Make amplitude plots

```{r ampPlots}
# "Mean and bootstrap CI based on fixed-effects uncertainty ONLY"
facLevs = c("Rewarded 1st 50 Buzzes\n (buzz #51-100)", 
               "Rewarded 2nd 50 Buzzes\n (buzz #1-50)", 
               "Rewarded 1st 50 Buzzes\n (buzz #1-50)",
               "Rewarded 2nd 50 Buzzes\n (buzz #50-100)"
               )

pframe$t2 <- factor(facLevs,  levels = facLevs[c(3,1,2,4)])


g00 <- ggplot(pframe, aes(x=treatment_rewarded, color = treatment_rewarded, y=predMean))+
     geom_point()+ 
     labs(y = expression ("Sonication amplitude "(m~s^{-2})), x = "Bee Rewarded") + 
    theme(legend.position = "none", 
          legend.background = element_rect(fill=alpha('gray90', 0.5))) +
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1) + 
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded?", begin =0.3, end = 0.8)
g00
ggsave(plot = g00, filename = file.path(figDir, "StartStopPreds_amp2.pdf"), width =5, height = 3.5)


fac2Levs = c("51-100","1-50", "1-50","51-100")
              

pframe$buzzNums <- factor(fac2Levs,  levels = fac2Levs[c(3,1)])
pframe$beesRewarded = mapvalues(pframe$treatment_rewarded, from  = c(" F", " T"), 
                                to = c("No", "Yes"))

g001 <- ggplot(pframe, aes(x=buzzNums, color = beesRewarded, y=predMean))+
     geom_point()+ 
     labs(y =expression ("Sonication amplitude "(m~s^{-2})), x = "Sonication Number") + 
    theme(legend.position = "none", 
          legend.background = element_rect(fill=alpha('gray90', 0.5))) +
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1) + 
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded", begin =0.3, end = 0.8) +
  facet_wrap(~trt2)
g001
ggsave(plot = g001, filename = file.path(figDir, "StartStopPreds_amp3_mod0.pdf"), width =5, height = 3.5)


```


# plot raw data

```{r}

sl <- bees
head(sl)
aggdata <- aggregate(log(sl$amp_acc), by=list(colNum = sl$beeID, trt = sl$treatment, rewarded = sl$treatment_rewarded ), FUN=mean, na.rm=TRUE)
colnames(aggdata)[colnames(aggdata) == "x"] = "amp"
head(aggdata)


ggplot(aggdata, aes(x=rewarded, color = rewarded, y=amp))+
     geom_boxplot()+ 
     labs(y =expression ("Sonication amplitude "(m~s^{-2})), x = "Sonication Number > 50") + 
    theme(legend.position = "none", 
          legend.background = element_rect(fill=alpha('gray90', 0.5))) +
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded", begin =0.3, end = 0.8) +
  facet_wrap(~trt)
```


# bootstrap for a different model

```{r bootstrap_Amp2, cache = TRUE}
# set number of bootstrap replicates for models
nbootSims = 5000

table(bees$hive) # more trials from hive 3

# using hive 7, since it's the one with the most data

pframe <- data.frame(trt2 = rep(levels(droplevels(bees$trt2)), 2),
                     treatment_rewarded = rep(levels(bees$treatment_rewarded), each = 2),
                     IT_mm = ITmean, 
                     hive = factor(7, levels = levels(bees$hive)),  
                     beeID = 99999)
pframe$amp_acc <- 0

m00 <- update(m2, .~., REML = TRUE)
pp <- exp(predict(m00, newdata = pframe, re.form=NA, type = 'response')) # re.form sets all random effects to 0



### Calculate CI's (using bootstrap, not accounting for random effects)
bb2 <- bootMer(m00, FUN=function(x) predict(x, pframe, re.form=NA, type = 'response'), nsim = nbootSims)
print(paste("Number of bootstrap samples", nrow(bb2$t)))
bb2_se <-apply(bb2$t,2,function(x) quantile(x, probs = c(0.025, 0.975)))
pframe$blo<-exp(bb2_se[1,])
pframe$bhi<-exp(bb2_se[2,])
pframe$predMean <- pp
pframe
```


```{r}

pframe$beesRewarded = mapvalues(pframe$treatment_rewarded, from  = c(" F", " T"), 
                                to = c("No", "Yes"))

g001 <- ggplot(pframe, aes(x=beesRewarded, color = beesRewarded, y=predMean))+
     geom_point()+ 
     labs(y =expression ("Sonication amplitude "(m~s^{-2})), x = "Bee rewarded") + 
    theme(legend.position = "none", 
          legend.background = element_rect(fill=alpha('gray90', 0.5))) +
     geom_errorbar(aes(ymin = blo, ymax = bhi), width = 0.1) + 
  scale_color_viridis(discrete = TRUE, name  = "Bees Rewarded", begin =0.3, end = 0.8)
g001
ggsave(plot = g001, filename = file.path(figDir, "StartStopPreds_amp3_finalMod.pdf"), width =5, height = 3.5)
```


#### Amplitude GAMM4

#### GAMM4 shows time series

```{r}

# start with gamm so I can show change by visit number
g01 = gamm4(log(amp_acc) ~ s(visitNum, by = trt2) + IT_mm + hive , family = "inverse.gaussian", random =  ~(1|beeID), data = bees)
par(mfrow = c(2,2))
aab <- plot(g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE, residuals = FALSE,pch = 20)
summary(g01$gam) # Summary for paper 
summary(g01$mer)
dev.off()


# save gamm plots
pdf(file.path(figDir, "Gamm_amp_startStop_1st50.pdf"), width = 4.5, height = 3.5)
par(mai= c(0.9,1,0.3,0.3))
dd = expression(paste('Estimated ', Delta, " log amplitude ", (m~s^{-2})))
plot.gam(select = 1, xlab = "Visit number", ylab = dd, g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE, bty = "l")
mtext("bees rewarded first", side = 2, padj = -3.5)
abline(v=50, lty = 2)
dev.off()


pdf(file.path(figDir, "Gamm_amp_startStop_2nd50.pdf"), width = 4.5, height = 3.5)
par(mai= c(0.9,1,0.3,0.3))
dd = expression(paste('Estimated ', Delta, " log amplitude ", (m~s^{-2})))
plot.gam(select = 2, xlab = "Visit number", ylab = dd, g01$gam, all.terms = TRUE, rug = FALSE, shade = TRUE, bty = "l")
mtext("bees rewarded second", side = 2, padj = -3.5)
dev.off()
```
