---
title: "Visualize experiment from hives 7-8"
author: "Callin Switzer"
date: "28 June 2017"
output:
  pdf_document: default
  html_document: default
---

<pre>
## Visualize experiment from hives 7-8
## Bees were given two treatments -- reward first or second
</pre>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/callinswitzer/Dropbox/SonicationLearningManuscript")
setwd("/Users/callinswitzer/Dropbox/SonicationLearningManuscript")
```

# Install packages
```{r, warning=FALSE, message = FALSE}
ipak <- function(pkg){
     new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
     if(length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
     sapply(pkg, require, character.only = TRUE)
}

packages <- c("ggplot2", "scales", "multcomp", "plyr", "car", "lme4", "signal", "viridis", "survival", "GGally", "survminer")
ipak(packages)


# print session info and time the code was run last
Sys.time()
sessionInfo()
```


# Read in data
```{r}
bees <- read.csv("freqLearn_Hive78_long.csv")
head(bees)

# add metadata to dataframe
splitMetadata <- strsplit(as.character(bees$BeeNumCol), split = "_" )
metad1 <- data.frame(matrix(unlist(splitMetadata), nrow=length(splitMetadata), byrow=T),stringsAsFactors=FALSE)

metad1[, c(1, 6)] <- NULL

colnames(metad1) <- c("beeNum", "year", "month", "day", "hive", "trt")

bees2 <- cbind(bees, metad1)

head(bees2)
nrow(bees2)

length(unique(bees2$BeeNumCol))

```

## visualize data
```{r}
theme_set(theme_bw())
ggplot(bees2, aes(x = index, y = freq, color = trt)) +
     geom_point(alpha = 0.1) + 
     stat_smooth(method = "loess", span = 0.75)

bees2$treatment <- mapvalues(bees2$trt, from = c("RewFir", "RewSec"), to = c("Rewarded_1st_50_Buzzes", "Rewarded_2nd_50_Buzzes"))

ggplot(bees2, aes(x = index, y = freq, color = rewTF)) +
     stat_smooth(method = "loess", span = 1) + 
     xlab("Buzz Number") + 
     ylab("Buzz Frequency(hz)") + 
     facet_wrap(~treatment) + 
     scale_color_viridis(name = "Rewarded Buzzes", 
                         discrete = TRUE, begin = 0.2, end = 0.6)
ggsave("beeRewards.png", width = 8, height = 5)


# make survival plot
survBees <- t(sapply(unique(bees2$beeNum), function(x) {
     tmp <- bees2[bees2$beeNum == x, ]
     return(tmp[which.max(tmp$index), ])
}))


lastObs <- as.data.frame(survBees)
lastObs$index <- as.numeric(lastObs$index)


# 1 = censored, 2 = stopped buzzing for 5 min
lastObs$status <- mapvalues(lastObs$index == 100, from = c(TRUE, FALSE), to = c(1, 2))

# create survival object
lastObs$survObj <- with(lastObs, Surv(index, status == 2))
head(lastObs)

lastObs$treatment <- as.factor(unlist(lastObs$treatment))


## Kaplan-Meier estimator. 
km.as.one <- survfit(survObj ~ 1, data = lastObs, conf.type = "log-log")
km.by.trt <- survfit(survObj ~ treatment, data = lastObs, conf.type = "log-log")

par(mfrow = c(1,1))
plot(km.by.trt)

gsv <- ggsurvplot(km.by.trt, CI = FALSE) 

gsv$plot + facet_wrap(~treatment) + 
     theme_bw()   + 
     theme(legend.position = "none")  + 
     scale_color_viridis(discrete = TRUE, begin = 0.3, end = 0.7, option = "magma")  + 
     xlab("Buzz Number") + 
     ylab("Proportion Sonicating") 

ggsave("beeSonicationPercentage.png", width = 8, height = 5)


ggsurv(km.by.trt, CI = FALSE, surv.col = c("#a6cee3", "#1f78b4")) + 
     ylim(c(0, 1)) + 
     xlab("Buzz Number") + 
     ylab("Proportion Sonicating") 

xtabs(~lastObs$treatment + as.character(lastObs$hive))



ggplot(bees2, aes(x = as.factor(index), y = freq, fill = trt)) +
     geom_boxplot(outlier.color = NA, position = position_dodge(width = 1))
     
library(gridExtra)
MinMeanSEMMax <- function(x) {
  v <- c(mean(x) - sd(x)/sqrt(length(x)), mean(x) - sd(x)/sqrt(length(x)), mean(x), mean(x) + sd(x)/sqrt(length(x)), mean(x) + sd(x)/sqrt(length(x)))
  names(v) <- c("ymin", "lower", "middle", "upper", "ymax")
  v
}

g2 <- ggplot(bees2, aes(x = as.factor(index), y = freq, fill = trt)) +
  stat_summary(fun.data=MinMeanSEMMax, geom="boxplot", colour="red") + 
     facet_wrap(~trt) 
g2


par(mfrow = c(1,2))
plot(bees2$index[bees2$trt == "RewFir" ], bees2$freq[bees2$trt == "RewFir" ], 
     pch = 20, col = rgb(0,0,0,0.1))
smth = lowess(bees2$index[bees2$trt == "RewFir" ], bees2$freq[bees2$trt == "RewFir" ])
lines(smth, col = 'red', lwd = 5)

plot(bees2$index[bees2$trt == "RewSec" ], bees2$freq[bees2$trt == "RewSec" ], 
     pch = 20, col = rgb(0,0,0,0.1))
smth = lowess(bees2$index[bees2$trt == "RewSec" ], bees2$freq[bees2$trt == "RewSec" ])
lines(smth, col = 'blue', lwd = 5)
```


# visualize averages
```{r}
bees2$lessThan50 <- bees2$index <= 50


ggplot(bees2, aes(x = lessThan50, y = freq, fill = trt)) +
     geom_boxplot() 

MinMeanSEMMax2 <- function(x) {
  v <- c(min(x), mean(x) - sd(x)/sqrt(length(x)), mean(x), mean(x) + sd(x)/sqrt(length(x)), max(x))
  names(v) <- c("ymin", "lower", "middle", "upper", "ymax")
  v
}


g3 <- ggplot(bees2, aes(x = lessThan50, y = freq, fill = trt)) +
  stat_summary(fun.data=MinMeanSEMMax, geom="boxplot", colour="red") + 
     facet_wrap(~trt) 
g3


```


# calculate average for each individual before and after treatment
```{r}
beeAvgs <- as.data.frame(tapply(X = bees2$freq, INDEX = list(bees2$beeNum, bees2$lessThan50), mean))
beetrt <- as.data.frame(sapply(X = unique(bees2$beeNum), FUN = function(x) bees2[bees2$beeNum == x, ][1, "trt"]))

beeAvgs <- cbind(beeAvgs, beetrt)

colnames(beeAvgs) = c("Second50", "First50", "trt")

beeAvgs$beeNum = row.names(beeAvgs)

# convert to long format
library(tidyr)

data_long <- gather(beeAvgs, condition, frequency, First50, Second50)
data_long

dl <- data_long[order(data_long$beeNum), ]

head(dl)


g4 <- ggplot(dl, aes(x = condition, y = frequency, fill = trt)) +
     geom_boxplot() + 
     facet_wrap(~trt) + 
     geom_point(position = position_jitter(width = 0.1))
g4

g5 <- ggplot(dl, aes(x = condition, y = frequency, fill = trt)) +
  stat_summary(fun.data=MinMeanSEMMax, geom="boxplot", colour="red") + 
     facet_wrap(~trt) + 
     geom_point(position = position_jitter(width = 0.1))
g5


```


# lmer analysis

```{r}
head(bees2)


mod1 <- lmer(freq ~ index*trt + bees2$lessThan50*trt +  (1|beeNum), data = bees2 )
summary(mod1)

plot(mod1)

preds <- predict(mod1, re.form = NA)

plot(bees2$index, preds, col = as.factor(bees2$trt))

## gam
library(mgcv)

bees3 <- bees2[bees2$index < 50, ]
bees3 <- bees2
gam1 <- gam(bees3$freq ~ s(bees3$index, by = as.factor(bees3$trt)) + bees3$lessThan50 +
                 s(as.factor(bees3$beeNum),bs = "re"))
summary(gam1)


par(mfrow = c(2,2))
plot(gam1, residuals = FALSE, cex = 1)

# get predictions, setting rand. effect to 0
preds  = predict(gam1, exclude="s(as.factor(bees3$beeNum))")


plot(bees3$index, preds)

plot(y = residuals(gam1), bees2$index)

```


# Bootstrap (probably too simple to use for publication)
```{r}
head(beeAvgs)


beeAvgs$difference = beeAvgs$Second50 - beeAvgs$First50

ggplot(beeAvgs, aes(y = difference, x = trt)) + 
     geom_boxplot()


bootFun <- function(o){
     beeSamp <- beeAvgs[sample(1:nrow(beeAvgs), replace = TRUE), ]
     mns <- tapply(beeSamp$difference, INDEX = beeSamp$trt, mean, na.rm = TRUE)
     return(diff(mns))
}
 

bootDiffs <- replicate(n = 100000, bootFun())    

par(mfrow = c(1,1))
hist(bootDiffs)
abline(v = 0)

mean(bootDiffs > 0)
```


# bootstrap different time steps and creat 95% CI for the mean for each group
```{r, eval = FALSE}
head(bees2)


# for each index, make bootstrap 95% CI's
indxFun <- function(idx) {
     beeSmp <- bees2[bees2$index == idx, ]
     beeSmp2 <- beeSmp[sample(1:nrow(beeSmp), replace = TRUE), ]
     mns1 <- tapply(beeSmp2$freq, INDEX = beeSmp2$trt, mean, na.rm = TRUE)
     return(mns1)
}

bootF2 <- function(o) t(sapply(1:100, indxFun))

bootSamps <- replicate(n = 10000, bootF2())

df1 <- as.data.frame(bootSamps)

rewFirDF <- df1[, grep(pattern = "RewFir", colnames(df1))]
rewSecDF <- df1[, grep(pattern = "RewSec", colnames(df1))]

frQuantiles <- t(apply(rewFirDF, MARGIN = 1, FUN = quantile, c(0.025, 0.975)))
seQuantiles <- t(apply(rewSecDF, MARGIN = 1, FUN = quantile, c(0.025, 0.975)))

plot(x = 1:100, y = frQuantiles[, 1], type = 'l', ylim = c(300, 400))
lines(x = 1:100, y = frQuantiles[, 2], type = 'l')

lines(x = 1:100, y = seQuantiles[, 1], type = 'l', col = 'red')
lines(x = 1:100, y = seQuantiles[, 2], type = 'l', col = 'red')

# calculate mean

meanVals <- t(sapply(1:100, function(idx){
     beeSmp <- bees2[bees2$index == idx, ]
     mns1 <- tapply(beeSmp$freq, INDEX = beeSmp$trt, mean, na.rm = TRUE)
     return(mns1)
}))

meanVals

lines(x = 1:100, y = meanVals[,1], lwd = 3)
lines(x = 1:100, y = meanVals[,2], lwd = 3, col = 'red')


plot(x = 1:100, y = meanVals[,2], lwd = 3, col = 'red', type = 'l')
lines(x = 1:100, y = meanVals[,1], lwd = 3)

```


# A different bootstrap analysis -- resampling whole time-series
```{r, eval = FALSE}
beesList <- unique(bees2$beeNum)

# for each index, make bootstrap 95% CI's
indxFun <- function(o) {
     beeListSample <- sample(beesList, replace = TRUE)
     bootDF <- (lapply(beeListSample, function(x) return(bees2[bees2$beeNum == x, ])))
     bootDF1 <- ldply(bootDF, data.frame)
     
     # calculate mean for each timepoint
     meanVals <- t(sapply(X = 1:100, function(idx){
          beeSmp <- bootDF1[bootDF1$index == idx, ]
          mns1 <- tapply(beeSmp$freq, INDEX = beeSmp$trt, mean, na.rm = TRUE)
          return(mns1)
          }))

     return(meanVals)
}

head(indxFun())

bootSamps <- lapply(1:1000, indxFun)

rewFirls <- lapply(bootSamps, function(x) x[,1])
rewSecLs <- lapply(bootSamps, function(x) x[,2])

dfir <- colMeans(do.call("rbind", rewFirls))
dsec <- colMeans(do.call("rbind", rewSecLs))

plot(x = 1:100, y = dfir, type = 'l')
lines(x = 1:100, y = dsec, type = 'l', col = 'red')
lines(x = 1:100, y = dfir, type = 'l')

dfirQuants <- apply(X = do.call("rbind", rewFirls), MARGIN = 2, quantile, c(0.025, 0.975))
dsec <- (do.call("rbind", rewSecLs))


# put survival plot in background

# get data about amount of pollen in different species

```




# amplitude analysis
```{r}
ggplot(bees2, aes(x = index, y = amp, color = trt)) +
     geom_point(alpha = 0.1) + 
     stat_smooth(method = "loess", span = 0.75)

bees2$treatment <- mapvalues(bees2$trt, from = c("RewFir", "RewSec"), to = c("Rewarded_1st_50_Buzzes", "Rewarded_2nd_50_Buzzes"))

ggplot(bees2, aes(x = index, y = amp, color = rewTF)) +
     stat_smooth(method = "loess", span = 1) + 
     xlab("Buzz Number") + 
     ylab("Buzz Amplitude (V)") + 
     facet_wrap(~treatment) + 
     scale_color_viridis(name = "Rewarded Buzzes", 
                         discrete = TRUE, begin = 0.2, end = 0.6)
ggsave("beeRewards_Amplitude.png", width = 8, height = 5)



```



